---
title: "Enhanced Forest Resources Inventory"
output:
  pdf_document: default
  html_document:
    df_print: paged
date: "2025-11-14"
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# **1** Introduction {-#introduction}
In recent years, many Canadian provinces have acquired airborne lidar data covering a large part of their territory. This technology makes it possible to redefine the description of forest attributes. Indeed, numerous initiatives have been carried out since the dissemination and democratization of lidar data and its derivative products.

## **1.1** Objective {-#introduction_objective}
The objective of this project is to automatically generate polygons that describe the forest. Both the delineation and description are done automatically through a series of steps that are described in the document. This project is a continuation of Ethan Berman's work at the University of British Columbia, which was conducted in 2023. Details are available on this page : https://bermane.github.io/ontario-inventory/index.html. The main improvement over previous work is that most of the steps have been integrated into a package, simplifying and standardizing them. The package is called eFRItools and is available here: https://github.com/metafor-ulaval/eFRItools. The functions can be used for other purposes, such as automated forest stand segmentation and forest attribute characterization.

## **1.3** Study area {-#introduction_studyarea}


## **1.3** Step {-#introduction_steps}
The steps for generating ecoforest polygons are as follows:
*Segmentation based on lidar metrics
*Extrapolation of forest composition variables using a k-nearest neighbor (kNN) imputation algorithm
*Construction of the attribute table to obtain the following attributes:
**LIST?

```{r Read library, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
library(terra)
library(tidyterra)
library(smoothr)
library(exactextractr)
library(rmapshaper)
library(sf)
library(pfif)
# remotes::install_github("metafor-ulaval/eFRItools")
library(eFRItools)
library(knitr)
library(basemapR)
```

```{r Parameters, include=FALSE, message=FALSE, warning=FALSE}
wd <- "D:/00_Ontario_eFRI/RMF" # wd
name <- "efri_polygon"
segmentation_wd <- paste0(wd, "/segmentations/", name)
dir.create(segmentation_wd)
epsg <- "EPSG:2958" # RMF
otb_dir <- "D:/00_Ontario_eFRI/logiciels/OTB-9.1.0-Win64/bin"
Sys.setenv(OTB_MAX_RAM_HINT = "65536") # 64 GO
Sys.setenv(OTB_MEMORY_AVAILABLE = "65536") # 64 GO
Sys.setenv(GDAL_CACHEMAX = "65536") # 64 GO

segmentation_metrics <- c("z_p95", "z_cv", "z_above2", "sagawi", "fractional_cover_05_2")

selected_dendrometrics <- c("Vmerch_ha", "qmdbh", "dens", "lor", "ba_ha")

best_models <- read.csv("D:/00_Ontario_eFRI/RMF/results/best_model.csv")

best_models %>%
  pull(knn_vars) %>% 
  strsplit(",") %>% 
  unlist() %>% 
  unique() -> best_models_variables

```

```{r Read metrics informations, include=FALSE, message=FALSE, warning=FALSE}
list.files(paste0(wd, "/metrics"), full.names = TRUE) %>%
  read_metrics() -> metrics_infos
```


```{r Read data, include=FALSE, message=FALSE, warning=FALSE}
# 1. Catalog
st_read(paste0(wd, "/ctg/ctg.shp")) %>% 
  st_transform(epsg) -> ctg

# Subset area in a 2 km radius circle in the centroid of area
ctg %>% 
  st_bbox() %>%
  st_as_sfc() %>%
  st_centroid() %>%
  st_buffer(5000) %>%
  st_bbox %>% 
  st_as_sfc() %>% 
  st_as_sf() -> subset_area

# 2. Metrics
metrics_infos %>%
  filter(name %in% c(best_models_variables, 
                     segmentation_metrics, 
                     selected_dendrometrics,
                     "z_p95", "z_above2", "fractional_cover_05_2", "slope", "sagawi")) %T>% 
  {pull(.,name) ->> metrics_names} %>% # Extract metrics names in the right order
  pull(path) %>% 
  map(rast) %>% 
  map(project, epsg, method = "bilinear") %>% 
  map(resample, .[[1]]) %>% 
  rast -> metrics

names(metrics) <- metrics_names # Assign metrics names

# 3. Masks
list(vect(paste0(wd, "/shp/roads.shp")),
     vect(paste0(wd, "/shp/waterbodies.shp"))) %>% 
  map(project, epsg) -> masks

# 4. Ecoforest polygons (fri)
st_read(paste0(wd, "/shp/PolygonForest.shp")) %>% 
  rowid_to_column("id") %>% 
  st_transform(epsg) -> fri_polygons

# 5. Landcover
rast(paste0(wd, "/metrics/other/landcover.tif")) %>%
      terra::project(epsg, method = "near") -> landcover

landcover_codes <- c(`1` = "Clear_Open_Water",
                     `2` = "Turbid_Water",
                     `3` = "Shoreline",
                     `4` = "Mudflats",
                     `5` = "Marsh",
                     `6` = "Swamp",
                     `7` = "Fen",
                     `8` = "Bog",
                     `10` = "Heath",
                     `11` = "Sparse_Treed",
                     `12` = "Treed_Upland",
                     `13` = "Deciduous_Treed",
                     `14` = "Mixed_Treed",
                     `15` = "Coniferous_Treed",
                     `16` = "Plantations_Treed_Cultivated",
                     `17` = "Hedge_Rows",
                     `18` = "Disturbance",
                     `19` = "Open_Cliff_Talus",
                     `20` = "Alvar",
                     `21` = "Sand_Barren_Dune",
                     `22` = "Open_Tallgrass_Prairie",
                     `23` = "Tallgrass_Savannah",
                     `24` = "Tallgrass_Woodland",
                     `25` = "Sand_Gravel_Mine_Tailings_Extraction",
                     `26` = "Bedrock",
                     `27` = "Communit_Infrastructure",
                     `28` = "Agriculture_Undifferentiated_Rural_Land_Use",
                     `157` = "Other",
                     `247` = "Cloud_Shadow")

# Attach levels (lookup table)
levels(landcover) <- data.frame(value = as.integer(names(landcover_codes)),
                                     class = unname(landcover_codes))

# 6. Forest age in 2019
rast(paste0(wd, "/metrics/other/forest_age_2019.tif")) %>%
      terra::project(epsg, method = "near") -> forest_age_2019

# 7. Forest fire between 1985-2020
rast(paste0(wd, "/metrics/other/forest_fire_1985_2020.tif")) %>%
      terra::project(epsg, method = "near") -> forest_fire_1985_2020

# 8. Forest harvest between 1985-2020
rast(paste0(wd, "/metrics/other/forest_harvest_1985_2020.tif")) %>%
      terra::project(epsg, method = "near") -> forest_harvest_1985_2020





# # CLIP ALL DATA FOR TESTING #
# ctg %<>% 
#   st_filter(subset_area)
# 
# metrics %<>% 
#   crop(subset_area)
# 
# masks %<>% 
#   map(crop, subset_area)
# 
# fri_polygons %<>% 
#   st_filter(subset_area)
# 
# landcover %<>% 
#   crop(subset_area)
# 
# forest_age_2019 %<>% 
#   crop(subset_area)
# 
# forest_fire_1985_2020 %<>% 
#   crop(subset_area)
# 
# forest_harvest_1985_2020 %<>% 
#   crop(subset_area)
```

# **2** Segmentation {-#segmentation}

The first step is to create polygon that will represent homogeneous forest stand. To perform this task, the Generic Region Merging algorithm from Orfeo ToolBox (OTB) is used.

```{r Segmentation, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
eFRI_segmentation(metrics = metrics[[segmentation_metrics]],
                  masks = masks,
                  thresh = 47,
                  spec = 0.6,
                  spat = 0.6,
                  method = "bs",
                  output_path = segmentation_wd,
                  output_name = "segmentation",
                  otb_dir = "D:/00_Ontario_eFRI/logiciels/OTB-9.1.0-Win64/bin") -> segmentation

segmentation %>% 
  st_write(paste0(segmentation_wd, "/data.gpkg"), 
           layer = "segmentation",
               quiet = T)
```

```{r Segmentation results, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  base_map(subset_area %>% st_transform(4326) %>% st_bbox(), 
           basemap = "google-satellite", increase_zoom = 3, nolabels = TRUE) +
  geom_sf(data = segmentation %>% st_geometry() %>% st_as_sf %>% st_filter(subset_area) %>% ms_clip(subset_area) %>% st_transform(4326), 
          color = "black", alpha = 0, size = 0.5) +
  theme_void()
```

## **2.3** Segmentation results {-#segmentation_results}

```{r Segmentation distribution results results, echo=FALSE, message=FALSE, warning=FALSE}
bind_rows(fri_polygons %>% st_geometry() %>% st_as_sf() %>% mutate(type = "fri"),
          segmentation %>% st_geometry() %>% st_as_sf() %>% mutate(type = "grm")) %>% 
  mutate(area = st_area(.) %>% as.numeric(),
         area_ha = area/10000,
         perimeter = st_perimeter(.) %>% as.numeric(),
         perimeter_area_ratio = perimeter/area,
         shape_index = perimeter/(2*sqrt(pi*area))) -> segmentation_results

segmentation_results %>% 
  ggplot() +
  geom_histogram(aes(x = shape_index,
                     y = stat(count / sum(count))*100,
                     color = type),
                 binwidth = 0.1,
                 fill="white", 
                 alpha=0.5, 
                 position="identity") +
  coord_cartesian(xlim = c(0, 10))

segmentation_results %>% 
  ggplot() +
  geom_histogram(aes(area_ha,
                     y = stat(count / sum(count))*100, 
                     color = type),
                 binwidth = 1,
                 fill="white", 
                 alpha=0.5, 
                 position="identity") +
  coord_cartesian(xlim = c(0, 30))
```

```{r Imputation, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
eFRI_imputation(segmentation = segmentation,
                segmentation_id_field = "id",
                forest_polygon = fri_polygons,
                metrics = metrics,
                landcover = landcover,
                forest_fire = forest_fire_1985_2020,
                forest_harvest = forest_harvest_1985_2020,
                ctg = ctg,
                lidar_year_field = "Fl_Cr_Y",
                forest_year_field = "YRUPD",
                forest_composition_field = "SPCOMP",
                forest_type_field = "POLYTYPE",
                target_var = best_models$target_var,
                knn_var = best_models$knn_vars) -> segmentation_imputation

segmentation_imputation %>% 
  st_write(paste0(segmentation_wd, "/data.gpkg"), 
           layer = "segmentation_imputation",
               quiet = T)
``` 

```{r eFRI attibute table, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
eFRI_attribute_table(segmentation = segmentation_imputation,
                     metrics = metrics,
                     selected_dendrometrics = selected_dendrometrics,
                     landcover = landcover,
                     forest_fire = forest_fire_1985_2020,
                     forest_harvest = forest_harvest_1985_2020,
                     forest_age = forest_age_2019) -> segmentation_data

segmentation_data %>% 
  st_write(paste0(segmentation_wd, "/data.gpkg"), 
           layer = "segmentation_data",
               quiet = T)
```
