---
title: "Enhanced Forest Resources Inventory"
output:
  pdf_document: default
  html_document:
    df_print: paged
date: "2025-11-14"
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# **1** Parameters {-#parameters}

## **1.1** Functions {-#parameters_functions}

First of all, functions that enable the computation of ecoforest polygon attributes are read. These are mostly functions that enable the reorganization of information in attribute tables.

```{r Read library, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
library(terra)
library(tidyterra)
library(smoothr)
library(exactextractr)
library(rmapshaper)
library(sf)
library(pfif)
remotes::install_github("metafor-ulaval/eFRItools")
library(eFRItools)
library(knitr)
library(basemapR)
```

```{r Set working directory, include=FALSE, message=FALSE, warning=FALSE}
"D:/00_Ontario_eFRI/RMF" -> wd
```

## **1.2** Metrics informations {-#parameters_metrics}

Next, all informations of available metrics are read. Three categories of metrics are used. First, metrics derived from raw lidar data, or from derivative products such as DTM, are named “lidar.” Second, dendrometric variables derived from lidar data and RandomForest models have been calculated. These metrics are intended to characterize certain forest attributes more explicitly. Third, data from spectral imagery are referred to as “sentinel_2.” This includes all available bands (11 bands) as well as a few indices calculated using these bands (10 indices). The following table shows all available metrics:

```{r Read metrics informations, include=FALSE, message=FALSE, warning=FALSE}
read_metrics(c("D:/00_Ontario_eFRI/RMF/metrics/lidar",
               "D:/00_Ontario_eFRI/RMF/metrics/sentinel2",
               "D:/00_Ontario_eFRI/RMF/metrics/dendro")) -> metrics_infos
```

```{r Show metricss list, echo=FALSE, message=FALSE, warning=FALSE}
knitr::kable(metrics_infos %>% dplyr::select(name, description, resolution, type), 
             caption = "Table 1: Metrics list")
```

## **1.3** Selected metrics {-#parameters_selected}

Next, the user must decide which metrics will be used to perform the segmentation of ecoforest polygons. The more metrics used, the smaller the resulting polygons will be. Using three to five metrics normally provides accurate segmentation. It is particularly important to select metrics that are as uncorrelated as possible and which have ecological significance. For example, if moisture is to be a discriminating factor in the segmentation, the saga wetness index should be used. Another example would be to use the fractional cover 50 cm - 2 m if the ecoforest stands are to be distinguished by their understory density. For now, the five metrics used allow for the segmentation of a wide variety of ecoforest stands. 

```{r Parameters, include=FALSE, message=FALSE, warning=FALSE}
epsg <- "EPSG:2958" # RMF
otb_dir <- "D:/00_Ontario_eFRI/logiciels/OTB-9.1.0-Win64/bin"
Sys.setenv(OTB_MAX_RAM_HINT = "65536") # 64 GO
Sys.setenv(OTB_MEMORY_AVAILABLE = "65536") # 64 GO
Sys.setenv(GDAL_CACHEMAX = "65536") # 64 GO

segmentation_metrics <- c("z_p95", "z_cv", "z_above2", "sagawi", "fractional_cover_05_2")

selected_dendrometrics <- c("Vmerch_ha", "qmdbh", "dens", "lor", "ba_ha")

best_models <- read.csv(paste0(wd, "/results/best_model.csv"))

best_models %>% 
  pull(knn_vars) %>% 
  strsplit(",") %>% 
  unlist() %>% 
  unique() -> best_models_variables

```

```{r Show parameters 1, echo=FALSE, message=FALSE, warning=FALSE}
segmentation_metrics
```

The user must also select the dendrometric variables that will be extracted in the final polygons. The median of the values in each of the polygons will be computed.

```{r Show parameters 2, echo=FALSE, message=FALSE, warning=FALSE}
selected_dendrometrics
```

Furthermore, a previous analysis allowed to identify the best combination of metrics for imputing variables characterizing forest species. Accuracy was assessed based on the ability of these metrics to correctly assign values to the original polygons. The four imputed variables are as follows: 

First leading species : "SP_NO_1"

Second leading species : "SP_NO_2"

Three functional group classification : "FUNCTIONAL_GROUP_3"

Five functional group classification : "FUNCTIONAL_GROUP_5"

```{r Show parameters 3, echo=FALSE, message=FALSE, warning=FALSE}
knitr::kable(best_models %>% dplyr::select(target_var, knn_vars, accuracy), 
             caption = "Table 2: Best metrics used for imputation")
```

## **1.4** Read data {-#parameters_data}

For the following steps, all base data are read. The list is as follows:

```{r Read data, include=FALSE, message=FALSE, warning=FALSE}
# 1. Catalog
st_read(paste0(wd, "/ctg/ctg.shp")) %>% 
  st_transform(epsg) -> ctg

# Subset area in a 2 km radius circle in the centroid of area
ctg %>% 
  st_bbox() %>%
  st_as_sfc() %>%
  st_centroid() %>%
  st_buffer(5000) %>%
  st_bbox %>% 
  st_as_sfc() %>% 
  st_as_sf() -> subset_area

# 2. Metrics
metrics_infos %>%
  filter(name %in% c(best_models_variables, 
                     segmentation_metrics, 
                     selected_dendrometrics,
                     "z_p95", "z_above2", "fractional_cover_05_2", "slope", "sagawi")) %T>% 
  {pull(.,name) ->> metrics_names} %>% # Extract metrics names in the right order
  pull(path) %>% 
  map(rast) %>% 
  map(project, epsg, method = "bilinear") %>% 
  map(resample, .[[1]]) %>% 
  rast -> metrics

names(metrics) <- metrics_names # Assign metrics names

# 3. Masks
list(vect(paste0(wd, "/shp/roads.shp")),
     vect(paste0(wd, "/shp/waterbodies.shp"))) %>% 
  map(project, epsg) -> masks

# 4. Ecoforest polygons (fri)
st_read(paste0(wd, "/shp/PolygonForest.shp")) %>% 
  rowid_to_column("id") %>% 
  st_transform(epsg) -> fri_polygons

# 5. Landcover
rast(paste0(wd, "/metrics/other/landcover.tif")) -> landcover

# 6. Forest age in 2019
rast(paste0(wd, "/metrics/other/forest_age_2019.tif")) -> forest_age_2019

# 7. Forest fire between 1985-2020
rast(paste0(wd, "/metrics/other/forest_fire_1985_2020.tif")) -> forest_fire_1985_2020

# 8. Forest harvest between 1985-2020
rast(paste0(wd, "/metrics/other/forest_harvest_1985_2020.tif")) -> forest_harvest_1985_2020
```

1. Index of LiDAR data in polygonal format. This layer allows the year of LiDAR data acquisition to be selected in order to adequately filter ecoforest polygons. The area in yellow will be used to display next data in order to properly see the details of the layers;

```{r Show data 3, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  geom_sf(data = ctg, aes(fill = Fl_Cr_Y), size = 0.5, alpha = 0.5) +
  geom_sf(data = subset_area, alpha = 0.2, color = "yellow") +
  labs(fill = "Acquisition year") +
  theme_void()
```


2. Metrics used for all steps, including:
      * Metrics used in the best imputation models;
      * Metrics selected for polygon segmentation;
      * Dendrometric variables selected for final extraction;
      * Basic metrics allowing the extraction of general stand attributes:
        * HEIGHT = z_p95;
        * CANOPY_COVER = z_above2;
        * DENSITY = fractional_cover_05_2;
        * SLOPE = slope;
        * MOISTURE = sagawi;

```{r Show data 1, echo=FALSE, message=FALSE, warning=FALSE}
for(metric in metrics_names){
  print(ggplot() +
          geom_spatraster(data = metrics[[metric]] %>% crop(subset_area)) +
          scale_fill_whitebox_c(palette = "muted") +
          theme_void() +
          labs(fill = metric))
}
```

3. Masks used for segmentation. Segmentation is therefore not performed on roads or waterbodies;

```{r Show data 2, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  base_map(subset_area %>% st_transform(4326) %>% st_bbox(), 
           basemap = "google-satellite", increase_zoom = 3, nolabels = TRUE) +
  geom_sf(data = masks[[1]] %>% st_as_sf %>% ms_clip(subset_area) %>% st_transform(4326), 
          color = "black", size = 1) +
  geom_sf(data = masks[[2]] %>% st_as_sf %>% ms_clip(subset_area) %>% st_transform(4326), 
          color = "blue", fill = "lightblue", size = 0.5) +
  theme_void()
```

4. Ecoforest polygons from the most recent inventory;

```{r Show data 4, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  base_map(subset_area %>% st_transform(4326) %>% st_bbox(), 
           basemap = "google-satellite", increase_zoom = 3, nolabels = TRUE) +
  geom_sf(data = fri_polygons %>% st_geometry() %>% st_as_sf %>% st_filter(subset_area) %>% ms_clip(subset_area) %>% st_transform(4326), 
          color = "black", alpha = 0, size = 0.5) +
  theme_void()
```

5. Ontario Land Cover Version 2.0;

```{r Show data 5, echo=FALSE, message=FALSE, warning=FALSE}
landcover_codes <- c(`1` = "Clear_Open_Water",
                     `2` = "Turbid_Water",
                     `3` = "Shoreline",
                     `4` = "Mudflats",
                     `5` = "Marsh",
                     `6` = "Swamp",
                     `7` = "Fen",
                     `8` = "Bog",
                     `10` = "Heath",
                     `11` = "Sparse_Treed",
                     `12` = "Treed_Upland",
                     `13` = "Deciduous_Treed",
                     `14` = "Mixed_Treed",
                     `15` = "Coniferous_Treed",
                     `16` = "Plantations_Treed_Cultivated",
                     `17` = "Hedge_Rows",
                     `18` = "Disturbance",
                     `19` = "Open_Cliff_Talus",
                     `20` = "Alvar",
                     `21` = "Sand_Barren_Dune",
                     `22` = "Open_Tallgrass_Prairie",
                     `23` = "Tallgrass_Savannah",
                     `24` = "Tallgrass_Woodland",
                     `25` = "Sand_Gravel_Mine_Tailings_Extraction",
                     `26` = "Bedrock",
                     `27` = "Communit_Infrastructure",
                     `28` = "Agriculture_Undifferentiated_Rural_Land_Use",
                     `157` = "Other",
                     `247` = "Cloud_Shadow")

landcover_plot <- landcover

# Attach levels (lookup table)
levels(landcover_plot) <- data.frame(value = as.integer(names(landcover_codes)),
                                     class = unname(landcover_codes))

# Remove NA cells from plotting
ggplot() +
  geom_spatraster(data = landcover_plot %>% crop(subset_area), na.rm = TRUE) +
  scale_fill_discrete(name = "Landcover", na.translate = FALSE) +
  theme_void()
```

6. Forest age in 2019 (CA Forest Age 2019 from url : https://opendata.nfis.org/mapserver/nfis-change_eng.html);

```{r Show data 6, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  geom_spatraster(data = forest_age_2019 %>% crop(subset_area), na.rm = TRUE) +
  scale_fill_whitebox_c(palette = "muted") +
  theme_void() +
  labs(fill = "Forest age (year)")
```

7. Forest fires between 1985 and 2020 (CA Forest Fires 1985-2020 from url : https://opendata.nfis.org/mapserver/nfis-change_eng.html);

```{r Show data 7, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  base_map(subset_area %>% st_transform(4326) %>% st_bbox(), 
           basemap = "google-satellite", increase_zoom = 3, nolabels = TRUE) +
  geom_spatraster(data = forest_fire_1985_2020 %>% crop(subset_area) %>% clamp(lower = 1985, values = F) %>% terra::project("EPSG:4326", method = "bilinear"), 
                  na.rm = TRUE) +
  scale_fill_whitebox_c(palette = "muted") +
  theme_void() +
  labs(fill = "Forest fire (year)")
```

8. Forest harvest between 1985 and 2020 (CA Forest Harvest 1985-2020 from url : https://opendata.nfis.org/mapserver/nfis-change_eng.html);

```{r Show data 8, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  base_map(subset_area %>% st_transform(4326) %>% st_bbox(), 
           basemap = "google-satellite", increase_zoom = 3, nolabels = TRUE) +
  geom_spatraster(data = forest_harvest_1985_2020 %>% crop(subset_area) %>% clamp(lower = 1985, values = F) %>% terra::project("EPSG:4326", method = "bilinear"), 
                  na.rm = TRUE) +
  scale_fill_whitebox_c(palette = "muted") +
  theme_void() +
  labs(fill = "Forest harvest (year)")
```

# **2** Segmentation {-#segmentation}
## **2.1** Generic region merging {-#segmentation_grm}

The first step after reading the data is to segment the selected metrics. To perform this task, the Generic Region Merging algorithm from Orfeo ToolBox (OTB) is used.

```{r Segmentation, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
eFRI_segmentation(metrics = rast(metrics),
                  masks = map(masks, vect),
                  thresh = 47,
                  spec = 0.6,
                  spat = 0.6,
                  method = "bs",
                  output_path = "D:/00_Ontario_eFRI/Test_Shiny",
                  output_name = "grm1",
                  otb_dir = "D:/00_Ontario_eFRI/logiciels/OTB-9.1.0-Win64/bin")
```

```{r Segmentation results, echo=FALSE, message=FALSE, warning=FALSE}
st_read(paste0(wd, "/segmentations/Rmarkdown/grm_polygons.gpkg"), 
        layer = "grm_polygons",
        quiet = T) -> grm_polygons

ggplot() +
  base_map(subset_area %>% st_transform(4326) %>% st_bbox(), 
           basemap = "google-satellite", increase_zoom = 3, nolabels = TRUE) +
  geom_sf(data = grm_polygons %>% st_geometry() %>% st_as_sf %>% st_filter(subset_area) %>% ms_clip(subset_area) %>% st_transform(4326), 
          color = "black", alpha = 0, size = 0.5) +
  theme_void()
```

## **2.2** Smoothing {-#segmentation_smoothing}

Since the resulting polygons were extracted from a raster with a resolution of 20 m, they have a “crumbly” appearance that follows the matrix. In order to smooth the polygons so that they appear more “natural,” the smooth function from the smoothr package was applied. The ksmooth method, which uses kernel smoothing, was used because it allows for smoother lines that are more similar to ecoforest polygons. Unfortunately, the function does not preserve the topology of polygons with adjacent boundaries, and each polygon is smoothed independently. This results in “silver polygons” between the smoothed polygons. This is therefore only a more interesting visual representation. For subsequent analyses, the raw polygons were used.

```{r Smoothing, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Smoothing
grm_polygons %>% 
  smoothr::smooth(method = "ksmooth", smoothness = 1) %>%
  st_make_valid() %>% 
  st_write(paste0(wd, "/segmentations/Rmarkdown/grm_polygons.gpkg"), 
           layer = "grm_polygons_smoothed",
           quiet = T)

```

```{r Segmentation smoothing results, echo=FALSE, message=FALSE, warning=FALSE}
st_read(paste0(wd, "/segmentations/Rmarkdown/grm_polygons.gpkg"), 
        layer = "grm_polygons_smoothed",
        quiet = T) -> grm_polygons_smoothed

ggplot() +
  base_map(subset_area %>% st_transform(4326) %>% st_bbox(), 
           basemap = "google-satellite", increase_zoom = 3, nolabels = TRUE) +
  geom_sf(data = grm_polygons_smoothed %>% st_geometry() %>% st_as_sf %>% st_filter(subset_area) %>% ms_clip(subset_area) %>% st_transform(4326), 
          color = "black", alpha = 0, size = 0.5) +
  theme_void()
```

## **2.3** Segmentation results {-#segmentation_results}

```{r Segmentation distribution results results, echo=FALSE, message=FALSE, warning=FALSE}
bind_rows(fri_polygons %>% st_geometry() %>% st_as_sf() %>% mutate(type = "fri"),
          grm_polygons %>% st_geometry() %>% st_as_sf() %>% mutate(type = "grm")) %>% 
  mutate(area = st_area(.) %>% as.numeric(),
         area_ha = area/10000,
         perimeter = st_perimeter(.) %>% as.numeric(),
         perimeter_area_ratio = perimeter/area,
         shape_index = perimeter/(2*sqrt(pi*area))) -> segmentation_results

segmentation_results %>% 
  ggplot() +
  geom_histogram(aes(x = shape_index,
                     y = stat(count / sum(count))*100,
                     color = type),
                 binwidth = 0.1,
                 fill="white", 
                 alpha=0.5, 
                 position="identity") +
  coord_cartesian(xlim = c(0, 10))

segmentation_results %>% 
  ggplot() +
  geom_histogram(aes(area_ha,
                     y = stat(count / sum(count))*100, 
                     color = type),
                 binwidth = 1,
                 fill="white", 
                 alpha=0.5, 
                 position="identity") +
  coord_cartesian(xlim = c(0, 30))
```

```{r Imputation, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
eFRI_imputation(segmentation = st_read("D:/00_Ontario_eFRI/Test_Shiny/grm1.shp"),
                forest_polygon = fri_polygons,
                metrics = rast(metrics),
                landcover = rast(landcover),
                forest_fire = rast(forest_fire_1985_2020),
                forest_harvest = rast(forest_harvest_1985_2020),
                ctg = st_read(paste0(wd, "/ctg/ctg.shp")) %>% st_transform("EPSG:2958"),
                lidar_year_field = "Fl_Cr_Y",
                forest_year_field = "YRUPD",
                forest_composition_field = "SPCOMP",
                forest_type_field = "POLYTYPE",
                target_var = best_models$target_var,
                knn_var = best_models$knn_vars)
```

```{r eFRI attibute table, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
eFRI_attribute_table(segmentation = st_read("D:/00_Ontario_eFRI/Test_Shiny/grm1.shp"),
                     metrics = rast(metrics),
                     landcover = rast(landcover),
                     forest_fire = rast(forest_fire_1985_2020),
                     forest_harvest = rast(forest_harvest_1985_2020),
                     forest_age = rast(forest_age_2019))

```
