---
title: "Enhanced Forest Resources Inventory"
output:
  pdf_document: default
  html_document:
    df_print: paged
date: "2025-11-14"
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Read library, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
library(terra)
library(tidyterra)
library(smoothr)
library(exactextractr)
library(rmapshaper)
library(sf)
library(pfif)
# remotes::install_github("metafor-ulaval/eFRItools")
library(eFRItools)
library(knitr)
library(basemapR)
```

# **1** Introduction {-#introduction}
In recent years, many Canadian provinces have acquired airborne lidar data covering a large part of their territory. This technology makes it possible to redefine the description of forest attributes. Indeed, numerous initiatives have been carried out since the dissemination and democratization of lidar data and its derivative products.

## **1.1** Objective {-#introduction_objective}
The objective of this project is to automatically generate polygons that describe the forest. Both the delineation and description are done automatically through a series of steps that are described in the document. This project is a continuation of Ethan Berman's work at the University of British Columbia, which was conducted in 2023. Details are available on this page : https://bermane.github.io/ontario-inventory/index.html. The main improvement over previous work is that most of the steps have been integrated into a package, simplifying and standardizing them. The package is called eFRItools and is available here: https://github.com/metafor-ulaval/eFRItools. The functions can be used for other purposes, such as automated forest stand segmentation and forest attribute characterization.

## **1.2** Study area {-#introduction_studyarea}


## **1.3** Step {-#introduction_steps}
The steps for generating ecoforest polygons are as follows:
*Segmentation based on lidar metrics
*Extrapolation of forest composition variables using a k-nearest neighbor (kNN) imputation algorithm
*Construction of the attribute table to obtain the following attributes:
**LIST?

```{r Parameters, include=FALSE, message=FALSE, warning=FALSE}
wd <- "D:/00_Ontario_eFRI/RMF" # wd
name <- "efri_polygon"
segmentation_wd <- paste0(wd, "/segmentations/", name)
dir.create(segmentation_wd)
epsg <- "EPSG:2958" # RMF
otb_dir <- "D:/00_Ontario_eFRI/logiciels/OTB-9.1.0-Win64/bin"
Sys.setenv(OTB_MAX_RAM_HINT = "65536") # 64 GO
Sys.setenv(OTB_MEMORY_AVAILABLE = "65536") # 64 GO
Sys.setenv(GDAL_CACHEMAX = "65536") # 64 GO

segmentation_metrics <- c("z_p95", "z_cv", "z_above2", "sagawi", "fractional_cover_05_2")

summary_metrics <- c("Vmerch_ha", "qmdbh", "dens", "lor", "ba_ha")

best_models <- read.csv("D:/00_Ontario_eFRI/RMF/results/best_model.csv")

best_models %>%
  pull(knn_vars) %>% 
  strsplit(",") %>% 
  unlist() %>% 
  unique() -> best_models_variables

```

```{r Read metrics informations, include=FALSE, message=FALSE, warning=FALSE}
list.files(paste0(wd, "/metrics"), full.names = TRUE) %>%
  read_metrics() -> metrics_infos
```

```{r Read data, include=FALSE, message=FALSE, warning=FALSE}
# 1. Catalog
st_read(paste0(wd, "/ctg/ctg.shp")) %>% 
  st_transform(epsg) -> ctg

# Subset area in a 2 km radius circle in the centroid of area
ctg %>% 
  st_bbox() %>%
  st_as_sfc() %>%
  st_centroid() %>%
  st_buffer(5000) %>%
  st_bbox %>% 
  st_as_sfc() %>% 
  st_as_sf() -> subset_area

# 2. Metrics
metrics_infos %>%
  filter(name %in% c(best_models_variables, 
                     segmentation_metrics, 
                     summary_metrics,
                     "z_p95", "z_above2", "fractional_cover_05_2", "slope", "sagawi")) -> metrics_infos_selected

metrics_infos_selected %>%
  pull(name) -> metrics_names

metrics_infos_selected %>%  
  pull(path) %>% 
  map(rast) %>% 
  map(project, epsg, method = "bilinear") %>% 
  map(resample, .[[1]]) %>% 
  rast -> metrics

names(metrics) <- metrics_names # Assign metrics names

# 3. Masks
list(vect(paste0(wd, "/shp/roads.shp")),
     vect(paste0(wd, "/shp/waterbodies.shp"))) %>% 
  map(project, epsg) -> masks

# 4. Ecoforest polygons (fri)
st_read(paste0(wd, "/shp/PolygonForest.shp")) %>% 
  rowid_to_column("id") %>% 
  st_transform(epsg) -> fri_polygons

# 5. Landcover
rast(paste0(wd, "/metrics/other/landcover.tif")) %>%
      terra::project(epsg, method = "near") -> landcover

landcover_codes <- c(`1` = "Clear_Open_Water",
                     `2` = "Turbid_Water",
                     `3` = "Shoreline",
                     `4` = "Mudflats",
                     `5` = "Marsh",
                     `6` = "Swamp",
                     `7` = "Fen",
                     `8` = "Bog",
                     `10` = "Heath",
                     `11` = "Sparse_Treed",
                     `12` = "Treed_Upland",
                     `13` = "Deciduous_Treed",
                     `14` = "Mixed_Treed",
                     `15` = "Coniferous_Treed",
                     `16` = "Plantations_Treed_Cultivated",
                     `17` = "Hedge_Rows",
                     `18` = "Disturbance",
                     `19` = "Open_Cliff_Talus",
                     `20` = "Alvar",
                     `21` = "Sand_Barren_Dune",
                     `22` = "Open_Tallgrass_Prairie",
                     `23` = "Tallgrass_Savannah",
                     `24` = "Tallgrass_Woodland",
                     `25` = "Sand_Gravel_Mine_Tailings_Extraction",
                     `26` = "Bedrock",
                     `27` = "Communit_Infrastructure",
                     `28` = "Agriculture_Undifferentiated_Rural_Land_Use",
                     `157` = "Other",
                     `247` = "Cloud_Shadow")

# Attach levels (lookup table)
levels(landcover) <- data.frame(value = as.integer(names(landcover_codes)),
                                     class = unname(landcover_codes))

# 6. Forest age in 2019
rast(paste0(wd, "/metrics/other/forest_age_2019.tif")) %>%
      terra::project(epsg, method = "near") -> forest_age_2019

# 7. Forest fire between 1985-2020
rast(paste0(wd, "/metrics/other/forest_fire_1985_2020.tif")) %>%
      terra::project(epsg, method = "near") -> forest_fire_1985_2020

# 8. Forest harvest between 1985-2020
rast(paste0(wd, "/metrics/other/forest_harvest_1985_2020.tif")) %>%
      terra::project(epsg, method = "near") -> forest_harvest_1985_2020





# CLIP ALL DATA FOR TESTING #
ctg %<>%
  st_filter(subset_area)

metrics %<>%
  crop(subset_area)

masks %<>%
  map(crop, subset_area)

fri_polygons %<>%
  st_filter(subset_area)

landcover %<>%
  crop(subset_area)

forest_age_2019 %<>%
  crop(subset_area)

forest_fire_1985_2020 %<>%
  crop(subset_area)

forest_harvest_1985_2020 %<>%
  crop(subset_area)
```

# **2** Segmentation {-#segmentation}

## **2.1** Segmentation basics {-#segmentation_basics}
The first step is to create polygon that will represent homogeneous forest stand. To perform this task, the Generic Region Merging tools from Orfeo ToolBox (OTB) is used. Three method can be used in this tools : Baatz & Schape, Eclidean Distance, and Full Lambda. For this work, the Baatz & Schape criterion method was used. This algorithm is the same used in the Multi-resolution Segmentation by the well known software eCognition and thus made it freely accessible. This algortihm works by merging iteratively small region together based on dissimilarity of spectral values. There is three parameters that control size and shape of resulting segments : threshold, weight of spectral homogeneity and weight of spatial homogeneity. Here is how these parameters influence segmentation:
*The threshold control maximum heterogeneity allowed when merging, controlling how fine or coarse the segmentation is :
**Small threshold → fine segmentation (many small regions)
**Large threshold → coarse segmentation (fewer, larger regions)

*The weight of spectral homogeneity gives to spectral similarity (pixel values, band means) when deciding whether regions should merge :
**If high → segmentation will mostly respect spectral values
**If low → spectral similarity matters little, other criteria (shape) dominate

*The weight of spatial homogeneity gives to shape similarity (compactness and smoothness) when deciding whether regions should merge :
**If high → the algorithm favors compact, smooth regions even if spectral similarity is weaker
**If low → region boundaries will mostly follow spectral homogeneity


If you need more details about parameters, eCognition made this [video](https://www.youtube.com/watch?v=tt53hoRAw9Q) to explain parameter selection for their Multi-resolution Segmentation algorithm.

## **2.2** Segmentation inputs {-#segmentation_inputs}
For this step, the following metrics are used for polygon segmentation :
*z_p95 → Height of 95th percentiles of returns above 1.3 m
*z_cv → Coefficient of variation of returns above 1.3 m
*z_above2 → Canopy cover of first returns above 2 m
*sagawi → Saga wetness index
*fractional_cover_05_2 → Number of returns between 0.5 m and 2 m divided by all returns below 2 m

```{r Show segmentation data 1, echo=FALSE, message=FALSE, warning=FALSE}
for(metric in segmentation_metrics){
  print(ggplot() +
          geom_spatraster(data = metrics[[metric]] %>% crop(subset_area)) +
          scale_fill_whitebox_c(palette = "muted") +
          theme_void() +
          labs(fill = metric))
}
```

These metrics are only one example of what can be used, but it is important to note that other metrics can be used for segmentation. The metrics used will reflect how the segmentation polygons are divided. Thus, if a spectral index such as NDVI is used, the polygons will tend to separate forest stands with contrasting spectral signatures from these indices. For effective segmentation, it is recommended to use metrics that provide different information. A correlation matrix or principal component analysis (PCA) can be used to see relationships between metrics.

An additional data used in segmentation are masks where segmentation is therefore not performed. These are typically bodies of water and roads. 

```{r Show segmentation data 2, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  base_map(subset_area %>% st_transform(4326) %>% st_bbox(), 
           basemap = "google-satellite", increase_zoom = 3, nolabels = TRUE) +
  geom_sf(data = masks[[1]] %>% st_as_sf %>% ms_clip(subset_area) %>% st_transform(4326), 
          color = "black", size = 1) +
  geom_sf(data = masks[[2]] %>% st_as_sf %>% ms_clip(subset_area) %>% st_transform(4326), 
          color = "blue", fill = "lightblue", size = 0.5) +
  theme_void()
```

```{r Segmentation, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
eFRI_segmentation(metrics = metrics[[segmentation_metrics]],
                  masks = masks,
                  thresh = 47,
                  spec = 0.6,
                  spat = 0.6,
                  method = "bs",
                  clean_nodata = TRUE,
                  output_path = segmentation_wd,
                  output_name = "segmentation",
                  otb_dir = "D:/00_Ontario_eFRI/logiciels/OTB-9.1.0-Win64/bin") -> segmentation

segmentation %>% 
  st_write(paste0(segmentation_wd, "/data.gpkg"), 
           layer = "segmentation",
               quiet = T)
```

## **2.3** Segmentation results {-#segmentation_results}
The segmentation result therefore produces polygons representing forest stands according to the selected metrics. These polygons can therefore be expected to represent homogeneous forest stands.

```{r Segmentation results, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  base_map(subset_area %>% st_transform(4326) %>% st_bbox(), 
           basemap = "google-satellite", increase_zoom = 3, nolabels = TRUE) +
  geom_sf(data = segmentation %>% 
            st_geometry() %>% 
            st_as_sf %>% 
            st_filter(subset_area) %>% 
            ms_clip(subset_area) %>% 
            st_transform(4326), 
          color = "black", alpha = 0, size = 0.5) +
  theme_void()
```

The polygons produced by the segmentation can be now compared to forest inventory ressources polygons. Histograms are used to assess whether the distribution of the size and shape of the segmentation polygons are equivalent to those of the forest inventory ressources polygons. Area in hectare are use for size and shape index, which is determine by this formula : [perimeter/(2*sqrt(pi*area))] is use to assess shape.

```{r Segmentation distribution results results, echo=FALSE, message=FALSE, warning=FALSE}
bind_rows(fri_polygons %>% st_geometry() %>% st_as_sf() %>% mutate(type = "fri"),
          segmentation %>% st_geometry() %>% st_as_sf() %>% mutate(type = "grm")) %>% 
  mutate(area = st_area(.) %>% as.numeric(),
         area_ha = area/10000,
         perimeter = st_perimeter(.) %>% as.numeric(),
         perimeter_area_ratio = perimeter/area,
         shape_index = perimeter/(2*sqrt(pi*area))) -> segmentation_results

segmentation_results %>% 
  ggplot() +
  geom_histogram(aes(x = shape_index,
                     y = stat(count / sum(count))*100,
                     color = type),
                 binwidth = 0.1,
                 fill="white", 
                 alpha=0.5, 
                 position="identity") +
  coord_cartesian(xlim = c(0, 10))

segmentation_results %>% 
  ggplot() +
  geom_histogram(aes(area_ha,
                     y = stat(count / sum(count))*100, 
                     color = type),
                 binwidth = 1,
                 fill="white", 
                 alpha=0.5, 
                 position="identity") +
  coord_cartesian(xlim = c(0, 30))
```

```{r Imputation, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
eFRI_imputation(segmentation = segmentation,
                segmentation_id_field = "id",
                forest_polygon = fri_polygons,
                metrics = metrics,
                landcover = landcover,
                forest_fire = forest_fire_1985_2020,
                forest_harvest = forest_harvest_1985_2020,
                ctg = ctg,
                lidar_year_field = "Fl_Cr_Y",
                forest_year_field = "YRUPD",
                forest_composition_field = "SPCOMP",
                forest_type_field = "POLYTYPE",
                target_var = best_models$target_var,
                knn_var = best_models$knn_vars) -> segmentation_imputation

segmentation_imputation %>% 
  st_write(paste0(segmentation_wd, "/data.gpkg"), 
           layer = "segmentation_imputation",
               quiet = T)
``` 

```{r eFRI attibute table, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
eFRI_attribute_table(segmentation = segmentation_imputation,
                     metrics = metrics,
                     summary_metrics = summary_metrics,
                     landcover = landcover,
                     forest_fire = forest_fire_1985_2020,
                     forest_harvest = forest_harvest_1985_2020,
                     forest_age = forest_age_2019) -> segmentation_data

segmentation_data %>% 
  st_write(paste0(segmentation_wd, "/data.gpkg"), 
           layer = "segmentation_data",
               quiet = T)
```
