---
title: "Enhanced Forest Resources Inventory"
output:
  pdf_document: default
  html_document:
    df_print: paged
date: "2025-09-04"
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# **1** Parameters {-#parameters}

## **1.1** Functions {-#parameters_functions}

First of all, functions that enable the computation of ecoforest polygon attributes are read. These are mostly functions that enable the reorganization of information in attribute tables.

```{r Read library, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
library(terra)
library(tidyterra)
library(smoothr)
library(exactextractr)
library(rmapshaper)
library(sf)
library(pfif)
library(knitr)
library(basemapR)
```

```{r Read functions, include=FALSE, message=FALSE, warning=FALSE}
# Functions
rstudioapi::getSourceEditorContext()$path %>% 
  dirname() %>% 
  paste0("/functions") %>% 
  list.files(full.names = T) %>% 
  map(source)
```

```{r Show functions list, echo=FALSE, message=FALSE, warning=FALSE}
# Functions
rstudioapi::getSourceEditorContext()$path %>% 
  dirname() %>% 
  paste0("/functions") %>% 
  list.files %>% 
  gsub(".R", "", .)
```

```{r Set working directory, include=FALSE, message=FALSE, warning=FALSE}
"D:/00_Ontario_eFRI/RMF" -> wd
```

## **1.2** Metrics informations {-#parameters_metrics}

Next, all informations of available metrics are read. Three categories of metrics are used. First, metrics derived from raw lidar data, or from derivative products such as DTM, are named “lidar.” Second, dendrometric variables derived from lidar data and RandomForest models have been calculated. These metrics are intended to characterize certain forest attributes more explicitly. Third, data from spectral imagery are referred to as “sentinel_2.” This includes all available bands (11 bands) as well as a few indices calculated using these bands (10 indices). The following table shows all available metrics:

```{r Read metrics informations, include=FALSE, message=FALSE, warning=FALSE}
# lidar
lidar_metrics_desc <- c("dem" = "digital elevation model",
                        "dem_5m" = "digital elevation model",
                        "slope" = "slope in percent",
                        "aspect" = "slope orientation in degrees",
                        "z_above2" = "canopy cover of first returns above 2 m",
                        "z_above5" = "canopy cover of first returns above 5 m",                        
                        "z_above10" = "canopy cover of first returns above 10 m", 
                        "z_above15" = "canopy cover of first returns above 15 m",
                        "z_mean" = "average height of returns above 1.3 m",
                        "z_cv" = "coefficient of variation of returns above 1.3 m",
                        "z_kurt" = "kurtosis  of returns above 1.3 m",
                        "z_max" = "maximum height of returns above 1.3 ms",
                        "z_p5" = "height of 5th percentiles of returns above 1.3 m",
                        "z_p10" = "height of 10th percentiles of returns above 1.3 m",
                        "z_p20" = "height of 20th percentiles of returns above 1.3 m",
                        "z_p30" = "height of 30th percentiles of returns above 1.3 m",
                        "z_p40" = "height of 40th percentiles of returns above 1.3 m",
                        "z_p50" = "height of 50th percentiles of returns above 1.3 m",
                        "z_p60" = "height of 60th percentiles of returns above 1.3 m",
                        "z_p70" = "height of 70th percentiles of returns above 1.3 m",
                        "z_p80" = "height of 80th percentiles of returns above 1.3 m",
                        "z_p90" = "height of 90th percentiles of returns above 1.3 m",
                        "z_p95" = "height of 95th percentiles of returns above 1.3 m",
                        "z_p99" = "height of 99th percentiles of returns above 1.3 m",
                        "rumple_index" = "rumple index of raw lidar data",
                        "z_qav" = "average square height of returns above 1.3 m",
                        "z_skew" = "skewness of returns above 1.3 m",
                        "z_sd" = "standard deviation of returns above 1.3 m",
                        "avg_p99" = "avg/p99",
                        "fractional_cover_05_2" = "number of returns between 0.5 m and 2 m divided by all returns below 2 m",
                        "fractional_cover_2_5" = "number of returns between 2 m and 5 m divided by all returns below 5 m",
                        "fractional_cover_5_10" = "number of returns between 5 m and 10 m divided by all returns below 10 m",
                        "relative_topographic_position_60m" = "relative topographic position with 3 pixels in x and y (60 m) from whiteboxtools",
                        "relative_topographic_position_100m" = "relative topographic position with 5 pixels in x and y (100 m) from whiteboxtools",
                        "relative_topographic_position_200m" = "relative topographic position with 10 pixels in x and y (200 m) from whiteboxtools",
                        "relative_topographic_position_400m" = "relative topographic position with 20 pixels in x and y (400 m) from whiteboxtools",
                        "relative_topographic_position_800m" = "relative topographic position with 40 pixels in x and y (800 m) from whiteboxtools",
                        "twi" = "topographic wetness index",
                        "twi_5m" = "topographic wetness index",
                        "sagawi" = "saga wetness index",
                        "sagawi_5m" = "saga wetness index")

tibble(path = list.files(paste0(wd, "/metrics/lidar"), pattern = "\\.tif$", full.names = TRUE)) %>% 
  mutate(name = list.files(paste0(wd, "/metrics/lidar"), pattern = "\\.tif$") %>% gsub(".tif", "", .),
         desc = lidar_metrics_desc[name],
         res = path %>% 
           map(rast) %>% 
           map(res) %>% 
           map(1) %>% 
           unlist(),
         type = "lidar") -> lidar_metrics_infos


# sentinel2
sentinel2_metrics_desc <- c("B2" = "Blue - 490 nm",
                            "B3" = "Green - 560 nm",
                            "B4" = "Red - 665 nm",
                            "B5" = "Vegetation Red Edge 1 (RE1) - 705 nm",
                            "B6" = "Vegetation Red Edge 2 (RE2) - 740 nm",
                            "B7sr" = "Surface Reflectance Vegetation Red Edge 3 - 783 nm",
                            "B8" = "Near Infrared (NIR) - 842 nm",
                            "B8sr" = "Surface Reflectance Near Infrared (NIRsr) - 842 nm",
                            "B8Asr" = "Surface Reflectance Narrow (Nsr) - 865 nm",
                            "B11" = "Shortwave Infrared 1 (SWIR1) - 1610 nm",
                            "B12" = "Shortwave Infrared 2 (SWIR2) - 2200 nm",
                            "NDVI" = "Normalized Difference Vegetation Index ((NIR - Red) / (NIR + Red))",
                            "GNDVI" = "Green Normalized Difference Vegetation Index ((NIR - Green) / (NIR + Green))",
                            "NDVIre" = "Normalized Difference Vegetation Index using Red Edge ((RE1 - Red) / (RE1 + Red))",
                            "RVI" = "Ratio Vegetation Index (NIR / Red)",
                            "DVI" = "Difference Vegetation Index (NIR - Red)",
                            "SAVI" = "Soil Adjusted Vegetation Index ((NIR - Red) / (NIR + Red + 0.5))*(1+0.5)",  
                            "EVI" = "Enhanced Vegetation Index 2.5*((NIR - Red) / (NIR + 6*Red - 7.5*Blue + 1))",
                            "GCI" = "Green Chlorophyll Index ((NIR / Green) - 1)",
                            "NDMI" = "Normalized Difference Moisture Index ((NIR - SWIR1) / (NIR + SWIR1))",  
                            "NDWI" = "Normalized Difference Water Index ((NIR - SWIR2) / (NIR + SWIR2))")

tibble(path = list.files(paste0(wd, "/metrics/sentinel2"), pattern = "\\.tif$", full.names = TRUE)) %>% 
  mutate(name = list.files(paste0(wd, "/metrics/sentinel2"), pattern = "\\.tif$") %>% gsub(".tif", "", .),
         desc = sentinel2_metrics_desc[name],
         res = path %>% 
           map(rast) %>% 
           map(res) %>% 
           map(1) %>% 
           unlist(),
         type = "sentinel_2") -> sentinel2_metrics_infos

# dendro
dendro_metrics_desc <- c("AGB_ha" = "above ground biomass normalized per hectare (t C/ha)",
                         "AGB_ha_FOR" = "above ground biomass normalized per hectare (t C/ha) / mask for RMF",
                         "ba_ha" = "basal area / tree cross sectional area (approximated as a circle) at breast height (1.3 m) (m²/ha)",
                         "ba_ha_FOR" = "basal area / tree cross sectional area (approximated as a circle) at breast height (1.3 m) (m²/ha) / mask for RMF",
                         "ba_ha_min20" = "basal area / tree cross sectional area (approximated as a circle) at breast height (1.3 m) (m²/ha) / min 20 ??????", # À élucider le min 20
                         "dens" = "stem density with DBH > 7.1 cm (stems/ha)",
                         "dens_FOR" = "tem density with DBH > 7.1 cm (stems/ha) / mask for RMF",
                         "lor" = "lorey’s height /	average tree height weighted by basal area (m)",
                         "lor_FOR" = "lorey’s height /	average tree height weighted by basal area (m) / mask for RMF",
                         "qmdbh" = "quadratic mean of diameter at breast height (cm)",
                         "qmdbh_FOR" = "quadratic mean of diameter at breast height (cm) / mask for RMF",
                         "top_height" = "maximum height (m)",
                         "top_height_FOR" = "maximum height (m) / mask for RMF",
                         "V_ha" = "total whole stem volume per hectare	(m³/ha)",
                         "V_ha_FOR" = "total whole stem volume per hectare	(m³/ha) / mask for RMF",
                         "Vmerch_ha" = "total merchantable volume normalized per hectare where stump height is set to 0.2 m and minimum to diameter to 10 cm (m³/ha)",
                         "Vmerch_ha_FOR" = "total merchantable volume normalized per hectare where stump height is set to 0.2 m and minimum to diameter to 10 cm (m³/ha) / mask for RMF")

tibble(path = list.files(paste0(wd, "/metrics/dendro"), pattern = "\\.tif$", full.names = TRUE)) %>% 
  mutate(name = list.files(paste0(wd, "/metrics/dendro"), pattern = "\\.tif$") %>% gsub(".tif", "", .), # CA NE MARCHERA PAS POUR OVF, RENOMMER PLUS SIMPLEMENT
         desc = dendro_metrics_desc[name],
         res = path %>% 
           map(rast) %>% 
           map(res) %>% 
           map(1) %>% 
           unlist(),
         type = "dendrometric") %>% 
  filter(!grepl("_FOR$", name)) -> dendro_metrics_infos

bind_rows(lidar_metrics_infos,
          sentinel2_metrics_infos,
          dendro_metrics_infos) -> metrics_infos

rm(dendro_metrics_desc,
   lidar_metrics_desc,
   sentinel2_metrics_desc,
   lidar_metrics_infos,
   sentinel2_metrics_infos,
   dendro_metrics_infos)
```

```{r Show metricss list, echo=FALSE, message=FALSE, warning=FALSE}
knitr::kable(metrics_infos %>% dplyr::select(name, desc, res, type), 
             caption = "Table 1: Metrics list")
```

## **1.3** Selected metrics {-#parameters_selected}

Next, the user must decide which metrics will be used to perform the segmentation of ecoforest polygons. The more metrics used, the smaller the resulting polygons will be. Using three to five metrics normally provides accurate segmentation. It is particularly important to select metrics that are as uncorrelated as possible and which have ecological significance. For example, if moisture is to be a discriminating factor in the segmentation, the saga wetness index should be used. Another example would be to use the fractional cover 50 cm - 2 m if the ecoforest stands are to be distinguished by their understory density. For now, the five metrics used allow for the segmentation of a wide variety of ecoforest stands. 

```{r Parameters, include=FALSE, message=FALSE, warning=FALSE}
epsg <- "EPSG:2958" # RMF
otb_dir <- "D:/00_Ontario_eFRI/logiciels/OTB-9.1.0-Win64/bin"
Sys.setenv(OTB_MAX_RAM_HINT = "65536") # 64 GO
Sys.setenv(OTB_MEMORY_AVAILABLE = "65536") # 64 GO
Sys.setenv(GDAL_CACHEMAX = "65536") # 64 GO

segmentation_metrics <- c("z_p95", "z_cv", "z_above2", "sagawi", "fractional_cover_05_2")

selected_dendrometrics <- c("Vmerch_ha", "qmdbh", "dens", "lor", "ba_ha")

best_models <- read.csv(paste0(wd, "/results/best_model.csv"))

best_models %>% 
  pull(knn_vars) %>% 
  strsplit(",") %>% 
  unlist() %>% 
  unique() -> best_models_variables

```

```{r Show parameters 1, echo=FALSE, message=FALSE, warning=FALSE}
segmentation_metrics
```

The user must also select the dendrometric variables that will be extracted in the final polygons. The median of the values in each of the polygons will be computed.

```{r Show parameters 2, echo=FALSE, message=FALSE, warning=FALSE}
selected_dendrometrics
```

Furthermore, a previous analysis allowed to identify the best combination of metrics for imputing variables characterizing forest species. Accuracy was assessed based on the ability of these metrics to correctly assign values to the original polygons. The four imputed variables are as follows: 

First leading species : "SP_NO_1"

Second leading species : "SP_NO_2"

Three functional group classification : "FUNCTIONAL_GROUP_3"

Five functional group classification : "FUNCTIONAL_GROUP_5"

```{r Show parameters 3, echo=FALSE, message=FALSE, warning=FALSE}
knitr::kable(best_models %>% dplyr::select(target_var, knn_vars, accuracy), 
             caption = "Table 2: Best metrics used for imputation")
```

## **1.4** Read data {-#parameters_data}

For the following steps, all base data are read. The list is as follows:

```{r Read data, include=FALSE, message=FALSE, warning=FALSE}
# 1. Catalog
st_read(paste0(wd, "/ctg/ctg.shp")) %>% 
  st_transform(epsg) -> ctg

# Subset area in a 2 km radius circle in the centroid of area
ctg %>% 
  st_bbox() %>%
  st_as_sfc() %>%
  st_centroid() %>%
  st_buffer(5000) %>%
  st_bbox %>% 
  st_as_sfc() %>% 
  st_as_sf() -> subset_area

# 2. Metrics
metrics_infos %>%
  filter(name %in% c(best_models_variables, 
                     segmentation_metrics, 
                     selected_dendrometrics,
                     "z_p95", "z_above2", "fractional_cover_05_2", "slope", "sagawi")) %T>% 
  {pull(.,name) ->> metrics_names} %>% # Extract metrics names in the right order
  pull(path) %>% 
  map(rast) %>% 
  map(project, epsg, method = "bilinear") %>% 
  map(resample, .[[1]]) %>% 
  rast -> metrics

names(metrics) <- metrics_names # Assign metrics names

# 3. Masks
list(vect(paste0(wd, "/shp/roads.shp")),
     vect(paste0(wd, "/shp/waterbodies.shp"))) %>% 
  map(project, epsg) -> masks

# 4. Ecoforest polygons (fri)
st_read(paste0(wd, "/shp/PolygonForest.shp")) %>% 
  rowid_to_column("id") %>% 
  st_transform(epsg) -> fri_polygons

# 5. Landcover
rast(paste0(wd, "/metrics/other/landcover.tif")) -> landcover

# 6. Forest age in 2019
rast(paste0(wd, "/metrics/other/forest_age_2019.tif")) -> forest_age_2019

# 7. Forest fire between 1985-2020
rast(paste0(wd, "/metrics/other/forest_fire_1985_2020.tif")) -> forest_fire_1985_2020

forest_fire_1985_2020 %>% 
  as.polygons() %>% 
  st_as_sf() %>% 
  dplyr::filter(.[[1]] != 0) %>% 
  st_cast("MULTIPOLYGON") %>% 
  st_cast("POLYGON") %>% 
  rename(year_fire = 1) %>% 
  st_transform(epsg) -> forest_fire_1985_2020_poly

# 8. Forest harvest between 1985-2020
rast(paste0(wd, "/metrics/other/forest_harvest_1985_2020.tif")) -> forest_harvest_1985_2020

forest_harvest_1985_2020 %>% 
  as.polygons() %>% 
  st_as_sf() %>% 
  dplyr::filter(.[[1]] != 0) %>% 
  st_cast("MULTIPOLYGON") %>% 
  st_cast("POLYGON") %>% 
  rename(year_harvest = 1) %>% 
  st_transform(epsg) -> forest_harvest_1985_2020_poly
```

1. Index of LiDAR data in polygonal format. This layer allows the year of LiDAR data acquisition to be selected in order to adequately filter ecoforest polygons. The area in yellow will be used to display next data in order to properly see the details of the layers;

```{r Show data 3, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  geom_sf(data = ctg, aes(fill = Fl_Cr_Y), size = 0.5, alpha = 0.5) +
  geom_sf(data = subset_area, alpha = 0.2, color = "yellow") +
  labs(fill = "Acquisition year") +
  theme_void()
```


2. Metrics used for all steps, including:
      * Metrics used in the best imputation models;
      * Metrics selected for polygon segmentation;
      * Dendrometric variables selected for final extraction;
      * Basic metrics allowing the extraction of general stand attributes:
        * HEIGHT = z_p95;
        * CANOPY_COVER = z_above2;
        * DENSITY = fractional_cover_05_2;
        * SLOPE = slope;
        * MOISTURE = sagawi;

```{r Show data 1, echo=FALSE, message=FALSE, warning=FALSE}
for(metric in metrics_names){
  print(ggplot() +
          geom_spatraster(data = metrics[[metric]] %>% crop(subset_area)) +
          scale_fill_whitebox_c(palette = "muted") +
          theme_void() +
          labs(fill = metric))
}
```

3. Masks used for segmentation. Segmentation is therefore not performed on roads or waterbodies;

```{r Show data 2, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  base_map(subset_area %>% st_transform(4326) %>% st_bbox(), 
           basemap = "google-satellite", increase_zoom = 3, nolabels = TRUE) +
  geom_sf(data = masks[[1]] %>% st_as_sf %>% ms_clip(subset_area) %>% st_transform(4326), 
          color = "black", size = 1) +
  geom_sf(data = masks[[2]] %>% st_as_sf %>% ms_clip(subset_area) %>% st_transform(4326), 
          color = "blue", fill = "lightblue", size = 0.5) +
  theme_void()
```

4. Ecoforest polygons from the most recent inventory;

```{r Show data 4, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  base_map(subset_area %>% st_transform(4326) %>% st_bbox(), 
           basemap = "google-satellite", increase_zoom = 3, nolabels = TRUE) +
  geom_sf(data = fri_polygons %>% st_geometry() %>% st_as_sf %>% st_filter(subset_area) %>% ms_clip(subset_area) %>% st_transform(4326), 
          color = "black", alpha = 0, size = 0.5) +
  theme_void()
```

5. Ontario Land Cover Version 2.0;

```{r Show data 5, echo=FALSE, message=FALSE, warning=FALSE}
landcover_codes <- c(`1` = "Clear_Open_Water",
                     `2` = "Turbid_Water",
                     `3` = "Shoreline",
                     `4` = "Mudflats",
                     `5` = "Marsh",
                     `6` = "Swamp",
                     `7` = "Fen",
                     `8` = "Bog",
                     `10` = "Heath",
                     `11` = "Sparse_Treed",
                     `12` = "Treed_Upland",
                     `13` = "Deciduous_Treed",
                     `14` = "Mixed_Treed",
                     `15` = "Coniferous_Treed",
                     `16` = "Plantations_Treed_Cultivated",
                     `17` = "Hedge_Rows",
                     `18` = "Disturbance",
                     `19` = "Open_Cliff_Talus",
                     `20` = "Alvar",
                     `21` = "Sand_Barren_Dune",
                     `22` = "Open_Tallgrass_Prairie",
                     `23` = "Tallgrass_Savannah",
                     `24` = "Tallgrass_Woodland",
                     `25` = "Sand_Gravel_Mine_Tailings_Extraction",
                     `26` = "Bedrock",
                     `27` = "Communit_Infrastructure",
                     `28` = "Agriculture_Undifferentiated_Rural_Land_Use",
                     `157` = "Other",
                     `247` = "Cloud_Shadow")

landcover_plot <- landcover

# Attach levels (lookup table)
levels(landcover_plot) <- data.frame(value = as.integer(names(landcover_codes)),
                                     class = unname(landcover_codes))

# Remove NA cells from plotting
ggplot() +
  geom_spatraster(data = landcover_plot %>% crop(subset_area), na.rm = TRUE) +
  scale_fill_discrete(name = "Landcover", na.translate = FALSE) +
  theme_void()
```

6. Forest age in 2019 (CA Forest Age 2019 from url : https://opendata.nfis.org/mapserver/nfis-change_eng.html);

```{r Show data 6, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  geom_spatraster(data = forest_age_2019 %>% crop(subset_area), na.rm = TRUE) +
  scale_fill_whitebox_c(palette = "muted") +
  theme_void() +
  labs(fill = "Forest age (year)")
```

7. Forest fires between 1985 and 2020 (CA Forest Fires 1985-2020 from url : https://opendata.nfis.org/mapserver/nfis-change_eng.html);

```{r Show data 7, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  base_map(subset_area %>% st_transform(4326) %>% st_bbox(), 
           basemap = "google-satellite", increase_zoom = 3, nolabels = TRUE) +
  geom_spatraster(data = forest_fire_1985_2020 %>% crop(subset_area) %>% clamp(lower = 1985, values = F) %>% terra::project("EPSG:4326", method = "bilinear"), 
                  na.rm = TRUE) +
  scale_fill_whitebox_c(palette = "muted") +
  theme_void() +
  labs(fill = "Forest fire (year)")
```

8. Forest harvest between 1985 and 2020 (CA Forest Harvest 1985-2020 from url : https://opendata.nfis.org/mapserver/nfis-change_eng.html);

```{r Show data 8, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  base_map(subset_area %>% st_transform(4326) %>% st_bbox(), 
           basemap = "google-satellite", increase_zoom = 3, nolabels = TRUE) +
  geom_spatraster(data = forest_harvest_1985_2020 %>% crop(subset_area) %>% clamp(lower = 1985, values = F) %>% terra::project("EPSG:4326", method = "bilinear"), 
                  na.rm = TRUE) +
  scale_fill_whitebox_c(palette = "muted") +
  theme_void() +
  labs(fill = "Forest harvest (year)")
```

# **2** Segmentation {-#segmentation}
## **2.1** Generic region merging {-#segmentation_grm}

The first step after reading the data is to segment the selected metrics. To perform this task, the Generic Region Merging algorithm from Orfeo ToolBox (OTB) is used.

```{r Segmentation, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Segmentation
metrics[[segmentation_metrics]] %>%
  pre_processing(masks = masks, smooth = 0) %>% # Normalizes the values to range in ⁠[0,255] with 1 and 99 percentile⁠
  generic_region_merging(thresh = 47, 
                         spec = 0.6, 
                         spat = 0.6, 
                         simplify = FALSE,
                         ofile = paste0(wd, "/segmentations/Rmarkdown/grm_raster.tif"), # Save the segmentation in raster
                         otb_dir = otb_dir) -> grm_polygons

# Multipart to single part polygons
grm_polygons %<>%
  st_cast("MULTIPOLYGON") %>% 
  st_cast("POLYGON") %>% 
  rename("id_segmentation" = 1) %>% 
  rowid_to_column("id") %>% # Compute id
  st_transform(epsg)
  
# Write polygons
grm_polygons %>%   
  st_write(paste0(wd, "/segmentations/Rmarkdown/grm_polygons.gpkg"), 
           layer = "grm_polygons",
           quiet = T)
```

```{r Segmentation results, echo=FALSE, message=FALSE, warning=FALSE}
st_read(paste0(wd, "/segmentations/Rmarkdown/grm_polygons.gpkg"), 
        layer = "grm_polygons",
        quiet = T) -> grm_polygons

ggplot() +
  base_map(subset_area %>% st_transform(4326) %>% st_bbox(), 
           basemap = "google-satellite", increase_zoom = 3, nolabels = TRUE) +
  geom_sf(data = grm_polygons %>% st_geometry() %>% st_as_sf %>% st_filter(subset_area) %>% ms_clip(subset_area) %>% st_transform(4326), 
          color = "black", alpha = 0, size = 0.5) +
  theme_void()
```

## **2.2** Smoothing {-#segmentation_smoothing}

Since the resulting polygons were extracted from a raster with a resolution of 20 m, they have a “crumbly” appearance that follows the matrix. In order to smooth the polygons so that they appear more “natural,” the smooth function from the smoothr package was applied. The ksmooth method, which uses kernel smoothing, was used because it allows for smoother lines that are more similar to ecoforest polygons. Unfortunately, the function does not preserve the topology of polygons with adjacent boundaries, and each polygon is smoothed independently. This results in “silver polygons” between the smoothed polygons. This is therefore only a more interesting visual representation. For subsequent analyses, the raw polygons were used.

```{r Smoothing, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Smoothing
grm_polygons %>% 
  smoothr::smooth(method = "ksmooth", smoothness = 1) %>%
  st_make_valid() %>% 
  st_write(paste0(wd, "/segmentations/Rmarkdown/grm_polygons.gpkg"), 
           layer = "grm_polygons_smoothed",
           quiet = T)

```

```{r Segmentation smoothing results, echo=FALSE, message=FALSE, warning=FALSE}
st_read(paste0(wd, "/segmentations/Rmarkdown/grm_polygons.gpkg"), 
        layer = "grm_polygons_smoothed",
        quiet = T) -> grm_polygons_smoothed

ggplot() +
  base_map(subset_area %>% st_transform(4326) %>% st_bbox(), 
           basemap = "google-satellite", increase_zoom = 3, nolabels = TRUE) +
  geom_sf(data = grm_polygons_smoothed %>% st_geometry() %>% st_as_sf %>% st_filter(subset_area) %>% ms_clip(subset_area) %>% st_transform(4326), 
          color = "black", alpha = 0, size = 0.5) +
  theme_void()
```

## **2.3** Segmentation results {-#segmentation_results}

```{r Segmentation distribution results results, echo=FALSE, message=FALSE, warning=FALSE}
bind_rows(fri_polygons %>% st_geometry() %>% st_as_sf() %>% mutate(type = "fri"),
          grm_polygons %>% st_geometry() %>% st_as_sf() %>% mutate(type = "grm")) %>% 
  mutate(area = st_area(.) %>% as.numeric(),
         area_ha = area/10000,
         perimeter = st_perimeter(.) %>% as.numeric(),
         perimeter_area_ratio = perimeter/area,
         shape_index = perimeter/(2*sqrt(pi*area))) -> segmentation_results

segmentation_results %>% 
  ggplot() +
  geom_histogram(aes(x = shape_index,
                     y = stat(count / sum(count))*100,
                     color = type),
                 binwidth = 0.1,
                 fill="white", 
                 alpha=0.5, 
                 position="identity") +
  coord_cartesian(xlim = c(0, 10))

segmentation_results %>% 
  ggplot() +
  geom_histogram(aes(area_ha,
                     y = stat(count / sum(count))*100, 
                     color = type),
                 binwidth = 1,
                 fill="white", 
                 alpha=0.5, 
                 position="identity") +
  coord_cartesian(xlim = c(0, 30))
```

```{r Compute values in grm polygons, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
grm_polygons %<>% 
  mutate(SOURCE = "eFRI") %>% # Source of data
  mutate(YRSOURCE = 2025) %>% # Year of the source
  mutate(area = st_area(.) %>% as.numeric()) %>% 
  mutate(perimeter = st_perimeter(.) %>% as.numeric()) %>% 
  mutate.metrics(metrics, fun = "median") %>% 
  mutate.centroid() %>% 
  mutate.landcover(landcover) %>% 
  mutate.simplify.landcover() %>% # Extract the landcover type and its proportion that cover the largest value of the polygon
  mutate.prop.forested() %>% 
  mutate.proportion(forest_fire_1985_2020, "fire") %>% 
  mutate.simplify.proportion("FIRE", 20) %>% # Extract the fire with the highest proportion : age (YRFIRE) and proportion (FIREPROP) (that cover at least 20 %)  
  mutate.proportion(forest_harvest_1985_2020, "harvest") %>% 
  mutate.simplify.proportion("HARVEST", 20) %>% # Extract the harvest with the highest proportion : age (YRHARVEST) and proportion (HARVESTPROP) (that cover at least 20 %)  
  mutate.age(forest_age_2019, 2019, 2025, "median")

# Extract all columns name that have a proportion of "HARVEST" or "FIRE"
grm_polygons %>% 
  st_drop_geometry() %>% 
  select(contains("HARVEST_") | contains("FIRE_")) %>% 
  names -> perturbation_column

# Mutate final values
grm_polygons %<>%
  mutate.origin.age(perturbation_column, 80) %>% # Extract the origin age of the stand (age of the last perturbation that cover more than 80 %)
  mutate(YRORG = ifelse(is.na(YRORG), (2025-age_median_2025), YRORG)) %>% # Assing the origin age if there was no perturbation
  mutate(AGE = 2025-YRORG) %>% # Recalculate the age based on the last perturbation that cover more than 80 %
  mutate.perturbation(perturbation_column, 20) %>% # Extrat the last perburtation with : age (YRDEP), type (DEPTYPE) and proportion (DEPPROP) (that cover at least 20 %)
  mutate(fractional_cover_05_2 = fractional_cover_05_2*100)


# # Compute statistics on dendrometrics variables
# grm_polygons %>% 
#   polygon_statistics(on = c("area", selected_dendrometrics)) %>% 
#   write.csv(paste0(wd, "/segmentations/Rmarkdown/grm_polygons_stats.csv"))
```

```{r Filter FRI polygon for imputation, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Find ids that had an harvest between inventory and lidar flight so that the inventory is not good
fri_polygons %>%
  dplyr::select(id, 
                year = YRUPD) %>% 
  st_intersection(ctg %>% dplyr::select(year_lidar = Fl_Cr_Y)) %>% 
  st_intersection(forest_harvest_1985_2020_poly %>% dplyr::select(year_harvest)) %>% 
  mutate(drop = ifelse(year_harvest > year & year_harvest < year_lidar, 1, 0)) %>% 
  filter(drop == 1) %>% 
  pull(id) %>% 
  unique -> drop_id_harvest

# Find ids that had a fire between inventory and lidar flight so that the inventory is not good
fri_polygons %>%
  dplyr::select(id, 
                year = YRUPD) %>% 
  st_intersection(ctg %>% dplyr::select(year_lidar = Fl_Cr_Y)) %>% 
  st_intersection(forest_fire_1985_2020_poly %>% dplyr::select(year_fire)) %>% 
  mutate(drop = ifelse(year_fire > year & year_fire < year_lidar, 1, 0)) %>% 
  filter(drop == 1) %>% 
  pull(id) %>% 
  unique -> drop_id_fire

# Remove polygons based on id
fri_polygons %<>% 
  filter(!id %in% c(drop_id_harvest, drop_id_fire))

# Extract metrics and variables for filtering
fri_polygons %<>% 
  mutate.metrics(metrics, fun = "median") %>%
  mutate.centroid() %>% 
  mutate.landcover(landcover) %>%  
  mutate.prop.forested()
  
# Keep forested poygons based on 3 data source
fri_polygons %<>% 
  filter(POLYTYPE == "FOR") %>% # keep polygon that are forested based on POLYTYPE
  filter(PROPFORESTED >= 50) %>% # keep polygon that have at least 50 % of forested area
  filter(z_p95 >= 5, z_above2 >= 50) # keep polygon with only high vegetation according to lidar data
```

L'impuitation devra être effectuée sur les meilleurs modèles, à modifier le code

```{r Imputation, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Compute imputation variables
fri_polygons %<>% 
  mutate.species.prop("SPCOMP") %>%
  mutate.species.order("SPCOMP") %>%
  mutate.forest.type %>% 
  mutate.functionals.groups

for(target_variable_selected in c("SP_NO_1", "SP_NO_2", "FUNCTIONAL_GROUP_3", "FUNCTIONAL_GROUP_5")){
  
  cat(paste0("Imputation of variable : ", target_variable_selected, "\n"))
  
  # Extract knn variables from the best model
  best_models %>% 
    filter(target_var == target_variable_selected) %>% 
    pull(knn_vars) %>%
    strsplit(",") %>% 
    unlist() -> knn_variables
  
  # Select only target columns of the fri polygons with values (drop_na)
  fri_polygons %>%
    dplyr::select(any_of(c("id", knn_variables, target_variable_selected))) %>% # The unique id, all variables used for knn imputation and the current target variable selected
    drop_na(everything()) -> fri_polygons_selected
  
  # Select only target columns of the grm polygons with values (drop_na)
  grm_polygons %>%
    dplyr::select(any_of(c("id", knn_variables))) %>% # The unique id and all variables used for knn imputation
    drop_na(everything()) -> grm_polygons_selected
  
  # Perform knn imputation
  knn.inputation(reference_polygons = fri_polygons_selected, # With all fri polygons with data (non-na)
                 target_polygons = grm_polygons_selected, # For all grm polygons with data (non-na)
                 knn_variables = knn_variables, # knn variables from the best model
                 target_variables = target_variable_selected, # For the current target variable selected
                 k = 5) %>%
    bind_cols(grm_polygons_selected, .) -> grm_polygons_selected
  
  # A left join is necessary to keep all polygons from original grm data (because some have na)
  grm_polygons_selected %>% 
    st_drop_geometry() %>% 
    dplyr::select(any_of(c("id", 
                           target_variable_selected))) %>%
    left_join(grm_polygons, ., by = "id") -> grm_polygons
}
```

```{r Final selection, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Order values ----
grm_polygons %>%
  dplyr::select(SOURCE,
                YRSOURCE,
                AREA = area,
                PERIMETER = perimeter,
                PROPFORESTED,
                POLYTYPE,
                POLYTYPEPROP,
                YRORG,
                AGE,
                YRDEP,
                DEPTYPE,
                DEPPROP,                  
                YRHARVEST,
                HARVESTPROP,
                YRFIRE,
                FIREPROP,
                selected_dendrometrics,
                HEIGHT = z_p95,
                CANOPY_COVER = z_above2,
                DENSITY = fractional_cover_05_2,
                SLOPE = slope,
                MOISTURE = sagawi,
                LEADSP = SP_NO_1,
                SECSP = SP_NO_2,
                FUNCTIONAL_GROUP_3, 
                FUNCTIONAL_GROUP_5) %>% 
  rename_with(toupper, selected_dendrometrics) %>% # A VOIR
  st_write(paste0(wd, "/segmentations/Rmarkdown/grm_polygons.gpkg"), 
           layer = "grm_polygons_data",
           quiet = T)
```