---
title: "Optimal Generic Region Merging parameters"
author: "Francis Lessard"
date: "2026-04-23"
output: pdf_document
---

```{r}
#| label = "setup",
#| include = FALSE
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#| label = "libraries",
#| echo = FALSE,
#| include = FALSE
library(tidyverse)
library(tidyterra)
library(future.apply)
library(future.callr)
library(future)
library(magrittr)
library(terra)
library(rmapshaper)
library(sf)
library(lasR)
library(lidR)
library(lidRmetrics)
library(plotly)
library(htmlwidgets)
library(eFRItools)
library(basemapR)
library(patchwork)
library(scatterplot3d)
```

```{r}
#| label = "function",
#| echo = FALSE,
#| include = FALSE
sample_farthest_point <- function(nb,
                                  res,
                                  sampling_area,
                                  seed) {

  # 1. Validate inputs
  if (!inherits(sampling_area, "sf")) {
    stop("Sampling_area must be sf objects.")
  }

  # 2. Prepare a permanent empty raster template (created ONCE)
  bbox <- sf::st_bbox(sampling_area)

  rast_template <- terra::rast(
    xmin  = bbox["xmin"],
    xmax  = bbox["xmax"],
    ymin  = bbox["ymin"],
    ymax  = bbox["ymax"],
    resolution = res,
    crs = sf::st_crs(sampling_area)$wkt
  )

  terra::values(rast_template) <- 1

  # 3. Select the first point randomly
  set.seed(seed)

  pt <- st_sample(sampling_area, 1, type = "random")

  # 4. Iteratively select farthest points
  for (i in 1:(nb-1)) {

    # Compute distance raster FROM currently selected points
    dist_raster <- terra::distance(rast_template, terra::vect(pt))

    # Mask raster
    dist_raster_masked <- terra::mask(dist_raster, sampling_area)

    # Plot results
    # plot(dist_raster_masked)
    # plot(pt, add = T, col = "red", pch = 16)

    # Locate farthest cell location within raster
    max_cell <- which.max(values(dist_raster_masked))
    farthest_cell <- rast(dist_raster_masked)
    values(farthest_cell) <- NA
    values(farthest_cell)[max_cell] <- 1

    # Randomly sample one point inside this cell
    farthest_cell_poly <- terra::as.polygons(farthest_cell)
    farthest_cell_poly <- sf::st_as_sf(farthest_cell_poly)
    set.seed(seed)
    new_pt <- st_sample(farthest_cell_poly, 1, type = "random")

    # Plot results
    # plot(new_pt, add = T, col = "blue", pch = 16)

    pt <- c(pt, new_pt)

  }

  return(pt)
}

smape <- function(actual, predicted) {
  (predicted - actual) / ((abs(actual) + abs(predicted)) / 2)
}
```

## Introduction

This document shows the step to find the optimals parameters of the Generic Region Merging (GRM) algorithm of the Orfeo ToolBox to automatically delineate operationally plausible forest stand. Photo-interpreted forest stands of the same area are the reference data. The optimization is performed on the Symmetric mean absolute percentage error (SMAPE) of polygon density, the overlapping kernel density of the polygon stand area and the overlapping kernel density of the polygon stand shape index. Best parameters were computed for two specific forest and five differents models, but performance metrics can be averaged to find best paramaters that suit the prefered scale.

```{r}
#| label = "base data",
#| eval = FALSE,
#| echo = FALSE,
#| include = FALSE
m <- c("z_p95", "z_above2", "z_cv", "z_skew", "sagawi", "B6", "B8", "B11")
inner_radius <- sqrt(10000^2*2)/2

for(f in c("RMF", "OVF")){

  st_read(paste0("D:/00_Ontario_eFRI/", f, "/shp/PolygonForest.shp"), quiet = T) -> polygon_forest

  st_read(paste0("D:/00_Ontario_eFRI/", f, "/ctg/ctg.shp"), quiet = T) %>%
    st_transform(st_crs(polygon_forest)) -> ctg

  ctg %>%
    st_geometry() %>%
    st_buffer(1000) %>%
    st_union() %>%
    st_buffer(-1000) -> area

  polygon_forest %>%
    filter(POLYTYPE == "FOR") %>%
    st_geometry() %>%
    st_as_sf() %>%
    sf::st_cast("MULTIPOLYGON") %>%
    st_cast("POLYGON") %>%
    st_filter(area) %>%
    st_buffer(500) %>%
    st_union() %>%
    st_buffer(-500) %>%
    st_buffer(-inner_radius)  %>%
    st_as_sf() -> sampling_inner_area

  sample_farthest_point(nb = 10,
                        res = 100,
                        sampling_area = sampling_inner_area,
                        seed = 1) %>%
    st_as_sf() -> sampling_point

  map_dfr(1:10,
          function(x){
            sampling_point[x,] %>%
              st_buffer(2500) %>%
              st_as_sf() %>%
              st_bbox() %>%
              st_as_sfc() %>%
              st_as_sf()
          }) %>%
    st_as_sf() -> sampling_square

  sampling_square %>%
    st_transform(st_crs(polygon_forest)) %>%
    st_write(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/base_data/sampling_square_", f, ".shp"))

  polygon_forest %>%
    st_intersection(sampling_square) %>%
    rowid_to_column("id") -> polygon_forest

  polygon_forest %>%
    st_write(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/base_data/polygon_forest_", f,".shp"))

  polygon_forest %>%
    sf::st_cast("MULTIPOLYGON") %>%
    st_cast("POLYGON") %>%
    filter(POLYTYPE == "FOR") -> polygon_forest_FOR

  polygon_forest %>%
    filter(!id %in% polygon_forest_FOR$id) %>%
    st_geometry() %>%
    sf::st_cast("MULTIPOLYGON") %>%
    st_cast("POLYGON") %>%
    st_as_sf() -> polygon_forest_NON_FOR

  list.files(paste0("D:/00_Ontario_eFRI/", f, "/metrics"), full.names = TRUE) %>%
    read_metrics() %>%
    filter(name %in% m) %T>%
    {pull(.,name) ->> metrics_names} %>% # Extract metrics names in the right order
    pull(path) %>%
    map(rast) %>%
    map(terra::project, paste0("EPSG:", st_crs(polygon_forest)$epsg), method = "bilinear") %>%
    map(terra::resample, .[[1]]) %>%
    rast -> metrics

  names(metrics) <- metrics_names

  metrics %<>%
    terra::crop(polygon_forest_FOR) %>%
    terra::mask(polygon_forest_FOR)

  # Save objects
  metrics %>%
    writeRaster(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/base_data/metrics_", f,".tif"))

  polygon_forest_FOR %>%
    st_write(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/base_data/polygon_forest_FOR_", f,".shp"))

  polygon_forest_NON_FOR %>%
    st_write(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/base_data/masks_", f,".shp"))

}
```

\newpage
## Sampling area

The two evaluated forest are located in the province of Ontario, in Canada. The first one is Romeo Malette Forest and are located in the middle of the province and is more conniferous. The second one is Ottawa Valley Forest and is located more south and that's why there is more deciduous forest. In each forest, ten smaller square area of 5 km x 5 km have been used to perform subsequent analysis. 

```{r}
#| label = "show base data 1",
#| echo = FALSE
for(f in c("RMF", "OVF")){
  st_read("D:/00_Ontario_eFRI/OVF/shp/PolygonForest.shp", quiet = T) -> polygon_forest
  
  st_read(paste0("D:/00_Ontario_eFRI/", f, "/ctg/ctg.shp"), quiet = T) %>%
    st_transform(st_crs(polygon_forest)) -> ctg

  ctg %>%
    st_geometry() %>%
    st_buffer(10) %>%
    st_union() %>%
    st_buffer(-10) %>% 
    st_transform(4326) -> area_clean
  
  area_clean %>% 
    st_bbox() -> bbox
  
  st_read(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/base_data/sampling_square_", f, ".shp"),
          quiet = T) %>% 
    st_transform(4326) -> sampling_square
  
  ggplot() +
    suppressMessages(base_map(bbox, basemap = "google-satellite", increase_zoom = 3, nolabels = T)) +
    geom_sf(data = sampling_square, fill = "#FFC107", alpha = 0.5) +
    geom_sf(data = area_clean, fill = NA, color = "white", lwd = 1) +    
    theme_bw() -> p
  
  assign(f, p)
}

RMF / OVF 
```

\newpage
To position theses area, only actual forest stand (POLYTYPE == "FOR") have been retain. 

```{r}
#| label = "show base data 2",
#| echo = FALSE
st_read("D:/00_Ontario_eFRI/OVF/shp/PolygonForest.shp", quiet = T) -> polygon_forest

st_read("D:/00_Ontario_eFRI/OVF/ctg/ctg.shp", quiet = T) %>%
  st_transform(st_crs(polygon_forest)) -> ctg

ctg %>%
  st_geometry() %>%
  st_buffer(10) %>%
  st_union() %>%
  st_buffer(-10) %>% 
  st_transform(4326) -> area_clean

area_clean %>% 
  st_bbox() -> bbox

ctg %>%
  st_geometry() %>%
  st_buffer(1000) %>%
  st_union() %>%
  st_buffer(-1000) -> area

polygon_forest %>% 
  filter(POLYTYPE == "FOR") %>%
  st_geometry() %>%
  st_as_sf() %>%
  sf::st_cast("MULTIPOLYGON") %>%
  st_cast("POLYGON") %>%
  st_filter(area) -> polygon_forest_temp

ggplot() +
    suppressMessages(base_map(bbox, basemap = "google-satellite", increase_zoom = 3, nolabels = T)) +
    geom_sf(data = polygon_forest_temp %>% st_transform(4326), color = NA, fill = "#D81B60", alpha = 0.5, lwd = 0) +
    geom_sf(data = area_clean, fill = NA, color = "white", lwd = 1) + 
    annotate("text", label = "map data © 2020 Google", size = 2) +
    theme_bw()
```

\newpage
A positive buffer of 500 m and then a negative one of 500 m again enable to merge close forest stand. 

```{r}
#| label = "show base data 3",
#| echo = FALSE
polygon_forest_temp %>% 
  st_buffer(500) %>%
  st_union() %>%
  st_buffer(-500) -> polygon_forest_temp

ggplot() +
  suppressMessages(base_map(bbox, basemap = "google-satellite", increase_zoom = 3, nolabels = T)) +
  geom_sf(data = polygon_forest_temp %>% st_transform(4326), color = NA, fill = "#D81B60", alpha = 0.5) +
  geom_sf(data = area_clean, fill = NA, color = "white", lwd = 1) +  
  theme_bw()
```

\newpage
A final inner buffer of approximately 7000 m was use to produce final polygon that allowed to place the center of each area. 

```{r}
#| label = "show base data 4",
#| echo = FALSE
inner_radius <- sqrt(10000^2*2)/2

polygon_forest_temp %>%
  st_buffer(-inner_radius)  %>%
  st_as_sf() -> polygon_forest_temp

ggplot() +
  suppressMessages(base_map(bbox, basemap = "google-satellite", increase_zoom = 3, nolabels = T)) +
  geom_sf(data = polygon_forest_temp %>% st_transform(4326), color = NA, fill = "#D81B60", alpha = 0.5) +
  geom_sf(data = area_clean, fill = NA, color = "white", lwd = 1) +  
  theme_bw()
```

\newpage
The first area was placed randomly and then the 9 next one were iteratively placed as far as possible of the last placed one. The next figure shows the last placed area in red and the distance raster that enable to fin the farthest position.

```{r}
#| label = "show base data 5",
#| echo = FALSE
sample_farthest_point(nb = 10,
                      res = 100,
                      sampling_area = polygon_forest_temp,
                      seed = 1) %>%
  st_as_sf() -> sampling_point

# Create raster
bbox_temp <- sf::st_bbox(polygon_forest_temp)

rast_template <- terra::rast(
  xmin  = bbox_temp["xmin"],
  xmax  = bbox_temp["xmax"],
  ymin  = bbox_temp["ymin"],
  ymax  = bbox_temp["ymax"],
  resolution = 100,
  crs = sf::st_crs(polygon_forest_temp)$wkt
)

terra::values(rast_template) <- 1

# Compute distance raster FROM currently selected points
dist_raster <- terra::distance(rast_template, terra::vect(sampling_point[1:9,]))

# Mask raster
dist_raster_masked <- terra::mask(dist_raster, polygon_forest_temp)

map_dfr(1:10,
        function(x){
          sampling_point[x,] %>%
            st_buffer(2500) %>%
            st_as_sf() %>%
            st_bbox() %>%
            st_as_sfc() %>%
            st_as_sf()
        }) %>%
  st_as_sf() -> sampling_square

suppressMessages(ggplot() +
  suppressMessages(base_map(bbox, basemap = "google-satellite", increase_zoom = 3, nolabels = T)) +
  geom_spatraster(data = (dist_raster_masked %>% terra::project("EPSG:4326", method = "bilinear"))/1000) +
  scale_fill_continuous(palette = "Viridis", na.value = NA) +
  geom_sf(data = sampling_square %>% st_transform(4326), fill = "#FFC107", alpha = 0.5) +
  geom_sf(data = sampling_point[1:9,] %>% st_transform(4326), pch = 16) +
  geom_sf(data = sampling_point[10,] %>% st_transform(4326), color = "red", pch = 16) +
  geom_sf(data = area_clean, fill = NA, color = "white", lwd = 1) +
  labs(fill = "Distance from first 9 points (km)") +
  theme(legend.position = "bottom") +
  theme_bw())
```

\newpage
For every area, only actual forest stand (POLYTYPE == "FOR") were used to perform segmentation.

```{r}
#| label = "show base data 6",
#| echo = FALSE
sampling_square %>% 
  slice(9) -> sampling_square_temp

sampling_square_temp %>% 
  st_transform(4326) %>% 
  st_bbox() -> bbox

sf::st_read("D:/00_Ontario_eFRI/segmentation/grm_parametrization/base_data/polygon_forest_FOR_OVF.shp", quiet = T) %>% 
  dplyr::select() %>%
  sf::st_cast("POLYGON") -> polygon_forest

polygon_forest %>% 
  st_filter(sampling_square_temp) -> polygon_forest_temp

ggplot() +
  suppressMessages(base_map(bbox, basemap = "google-satellite", increase_zoom = 3, nolabels = T)) +
  geom_sf(data = polygon_forest_temp %>% st_transform(4326), color = "black", fill = "#D81B60", alpha = 0.5) +
  theme_bw()
```

\newpage
## Segmentation

Five differents sets of metrics were used to perform segmentation :

- model_A : z_p95 + z_above2 + z_cv
- model_B : z_p95 + z_above2 + z_skew
- model_D : z_p95 + z_above2 + sagawi
- model_F : z_p95 + z_above2 + B6
- model_G : B6 + B8 + B11

The metrics used are shown in the follow figure.

```{r}
#| label = "show base data 7",
#| echo = FALSE
terra::rast("D:/00_Ontario_eFRI/segmentation/grm_parametrization/base_data/metrics_OVF.tif") -> metrics

metrics %>%
  terra::crop(sampling_square_temp %>% st_transform(st_crs(metrics))) %>%
  terra::mask(sampling_square_temp %>% st_transform(st_crs(metrics))) -> metrics_temp

map(names(metrics_temp),
    function(x){
      ggplot() +
        geom_spatraster(data = metrics_temp[[x]]) +
            scale_fill_whitebox_c(palette = "muted",
                          labels = scales::label_number(accuracy = 1),
                          breaks = scales::pretty_breaks(n = 5)) +
    theme_void() +
    labs(fill = x)
    }) -> f

f[[1]]
f[[2]]
f[[3]]
f[[4]]
f[[5]]
f[[6]]
f[[7]]  
f[[8]]

```

\newpage
## Parameters

Finally, the parameters tested are the following ones : 
1. The threshold control maximum heterogeneity allowed when merging, controlling how fine or coarse the segmentation is :
 - Small threshold → fine segmentation (many small regions)
 - Large threshold → coarse segmentation (fewer, larger regions)
 - This parameter was tested with values ranging from 10 to 100 in increments of 10

2. The weight of spectral homogeneity gives to spectral similarity (pixel values, band means) when deciding whether regions should merge :
 - If high → segmentation will mostly respect spectral values
 - If low → spectral similarity matters little, other criteria (shape) dominate
 - This parameter was tested with values ranging from 0.1 to 1 in increments of 0.1

3. The weight of spatial homogeneity gives to shape similarity (compactness and smoothness) when deciding whether regions should merge :
 - If high → the algorithm favors compact, smooth regions even if spectral similarity is weaker
 - If low → region boundaries will mostly follow spectral homogeneity
 - This parameter was tested with values ranging from 0.1 to 1 in increments of 0.1

Thus, a total of 1000 paramaters combinaisons was tested.

```{r}
#| label = "segmentation",
#| eval = FALSE,
#| echo = FALSE,
#| include = FALSE
plan(multisession, workers = round(ncores()*0.25))

forest <- c("RMF", "OVF")
thresh <- seq(10, 100, 10)
spec <- seq(0.1, 1, 0.1)
spat <- seq(0.1, 1, 0.1)
models <- list(model_A = c("z_p95", "z_above2", "z_cv"),
               model_B = c("z_p95", "z_above2", "z_skew"),
               model_D = c("z_p95", "z_above2", "sagawi"),
               model_F = c("z_p95", "z_above2", "B6"),
               model_G = c("B6", "B8", "B11"))

models_df <- data.frame(model_name = names(models),
                        model_var  = sapply(models, paste, collapse = ", "),
                        stringsAsFactors = FALSE)

params <- merge(expand.grid(forest = forest,
                            thresh = thresh,
                            spec = spec,
                            spat = spat,
                            stringsAsFactors = FALSE),
                models_df)

# Remove done files
params %>%
  mutate(done = file.exists(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/segmentation/", forest, "/", model_name, "/segmentation_thresh", thresh, "_spec", spec*100, "_spat", spat*100, "_smoothed.shp"))) %>%
  filter(!done) -> params

# Remove remaining raster files
list.files(path = "D:/00_Ontario_eFRI/segmentation/grm_parametrization/segmentation", full.names = T) %>%
  map(list.files, full.names = T) %>%
  unlist() %>%
  map(list.files, pattern = "\\.tif$", full.names = T) %>%
  unlist() %>%
  file.remove()

# params %>%
#   filter(forest == "RMF" & model_name == "model_F" & thresh == 60 & spec == 0.2 & spat == 0.1) -> param

# p <- 889
# x <- 1

future_lapply(seq_len(nrow(params)), function(p){

  library(eFRItools)
  library(sf)
  library(terra)
  library(stringr)

  param = params[p,]

  cat("Forest :", param$forest, "/ Model :", param$model_name, "/ Threshold :", param$thresh, "/ Spec :", param$spec, "/ Spat :", param$spat, "\n")

  if(!dir.exists(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/segmentation/", param$forest))){
    dir.create(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/segmentation/", param$forest))
  }

  if(!dir.exists(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/segmentation/", param$forest, "/", param$model_name))){
    dir.create(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/segmentation/", param$forest, "/", param$model_name))
  }

  output_name <- paste0("_thresh", param$thresh,
                        "_spec", param$spec*100,
                        "_spat", param$spat*100)

  st_read(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/base_data/sampling_square_", param$forest, ".shp"),
          quiet = T) -> sampling_square

  rast(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/base_data/metrics_", param$forest,".tif")) -> metrics

  st_read(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/base_data/masks_", param$forest,".shp"),
          quiet = T) -> masks

  map_dfr(seq_len(nrow(sampling_square)),
          function(x){

            metrics[[strsplit(param$model_var, ", ")[[1]]]] %>%
              terra::crop(sampling_square[x,]) %>%
              terra::mask(sampling_square[x,]) -> metrics_temp

            masks %>%
              sf::st_filter(sampling_square[x,]) -> masks_temp

            eFRI_segmentation(metrics = metrics_temp,
                              masks = list(vect(masks_temp)),
                              thresh = param$thresh,
                              spec = param$spec,
                              spat = param$spat,
                              method = "bs",
                              clean_nodata = TRUE,
                              output_path = paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/segmentation/", param$forest, "/", param$model_name),
                              output_name = paste0("segmentation", output_name, "_it", x),
                              otb_dir = "D:/00_Ontario_eFRI/logiciels/OTB-9.1.0-Win64/bin") %>%
              sf::st_geometry() %>%
              sf::st_as_sf() -> segmentation

            sf::st_geometry(segmentation) <- "geometry"

            return(segmentation)

          }) -> segmentation_merged

  list.files(path = paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/segmentation/", param$forest, "/", param$model_name),
             pattern = "\\.tif$",
             full.names = T) %>%
    stringr::str_subset(paste0("segmentation", output_name)) %>%
    file.remove()

  st_write(segmentation_merged,
           paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/segmentation/", param$forest, "/", param$model_name, "/segmentation", output_name, ".shp"),
           append = FALSE,
           quiet = T)

  chunk_size <- 1000
  segmentation_chunks <- split(segmentation_merged,
                               ceiling(seq_len(nrow(segmentation_merged)) / chunk_size))

  map_dfr(seq_along(segmentation_chunks),
          function(chunk_no) {

            cat("Processing chunk", chunk_no, "out of", length(segmentation_chunks), "\n")

            chunk <- segmentation_chunks[[chunk_no]]

            chunk %>%
              smoothr::smooth(method = "ksmooth", smoothness = 1) %>%
              sf::st_make_valid() -> chunk_smoothed

              return(chunk_smoothed)

          }) -> segmentation_smoothed

  segmentation_smoothed %>%
    st_cast("MULTIPOLYGON", warn = FALSE) %>%
    st_make_valid() -> segmentation_smoothed

  segmentation_smoothed[st_geometry_type(segmentation_smoothed) %in% c("POLYGON", "MULTIPOLYGON"),] -> segmentation_smoothed

  st_write(segmentation_smoothed,
           paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/segmentation/", param$forest, "/", param$model_name, "/segmentation", output_name, "_smoothed.shp"),
           append = FALSE,
           quiet = T)

  return(TRUE)

})
```

The next figure shows a example of a segmentation in the Ottawa Valley Forest for the model_A (z_p95 + z_above2 + z_cv) with mid parameters (threshold of 50, spectral homogeneity of 0.5 and spatial homogeneity of 0.5). There is also a example of a segmnetation that was smoothed to get a better shape index of each polygon forest stand.

```{r}
#| label = "show segmentation 1",
#| echo = FALSE
polygon_forest %>%
  mutate(area = as.numeric(sf::st_area(.)),
         perimeter = as.numeric(sf::st_perimeter(.)),
         perimeter_area_ratio = perimeter/area,
         shape_index = perimeter / sqrt(pi * area)) %>%
  filter_if(is.numeric, is.finite) -> polygon_forest_temp

suppressMessages(
  st_read("D:/00_Ontario_eFRI/segmentation/grm_parametrization/segmentation/OVF/model_A/segmentation_thresh50_spec50_spat50.shp",
        quiet = T) %>% 
  sf::st_make_valid() %>%
  sf::st_cast("MULTIPOLYGON") %>%
  sf::st_cast("POLYGON") %>%
  mutate(area = as.numeric(st_area(.)),
         perimeter = as.numeric(st_perimeter(.)),
         perimeter_area_ratio = perimeter/area,
         shape_index = perimeter / sqrt(pi * area)) %>%
  filter_if(is.numeric, is.finite)) -> segmentation 

suppressMessages(st_read("D:/00_Ontario_eFRI/segmentation/grm_parametrization/segmentation/OVF/model_A/segmentation_thresh50_spec50_spat50_smoothed.shp",
        quiet = T) %>% 
  sf::st_make_valid() %>%
  sf::st_cast("MULTIPOLYGON") %>%
  sf::st_cast("POLYGON") %>%
  mutate(area = as.numeric(st_area(.)),
         perimeter = as.numeric(st_perimeter(.)),
         perimeter_area_ratio = perimeter/area,
         shape_index = perimeter / sqrt(pi * area)) %>%
  filter_if(is.numeric, is.finite)) -> segmentation_smoothed

polygon_forest_temp %>%
  dplyr::select(area) %>%
  terra::rasterize(metrics, field = "area") %>%
  terra::values(na.rm = TRUE) -> polygon_forest_area_raster

polygon_forest_area_raster[is.finite(polygon_forest_area_raster)] -> polygon_forest_area_raster

segmentation %>%
  dplyr::select(area) %>%
  terra::rasterize(metrics, field = "area") %>%
  terra::values(na.rm = TRUE) -> segmentation_area_raster

segmentation_area_raster[is.finite(segmentation_area_raster)] -> segmentation_area_raster

polygon_forest_temp %>%
  dplyr::select(shape_index) %>%
  terra::rasterize(metrics, field = "shape_index") %>%
  terra::values(na.rm = TRUE) -> polygon_forest_shape_index_raster

polygon_forest_shape_index_raster[is.finite(polygon_forest_shape_index_raster)] -> polygon_forest_shape_index_raster

segmentation %>%
  dplyr::select(shape_index) %>%
  terra::rasterize(metrics, field = "shape_index") %>%
  terra::values(na.rm = TRUE) -> segmentation_shape_index_raster

segmentation_shape_index_raster[is.finite(segmentation_shape_index_raster)] -> segmentation_shape_index_raster

segmentation_smoothed %>%
  dplyr::select(shape_index) %>%
  terra::rasterize(metrics, field = "shape_index") %>%
  terra::values(na.rm = TRUE) -> segmentation_smoothed_shape_index_raster

segmentation_smoothed_shape_index_raster[is.finite(segmentation_smoothed_shape_index_raster)] -> segmentation_smoothed_shape_index_raster

ggplot() +
  suppressMessages(base_map(bbox, basemap = "google-satellite", increase_zoom = 3, nolabels = T)) +
  geom_sf(data = segmentation %>% st_filter(sampling_square_temp) %>% st_transform(4326), color = "black", fill = "#7A9A05", alpha = 0.5) +
  theme_bw() -> f1

ggplot() +
  suppressMessages(base_map(bbox, basemap = "google-satellite", increase_zoom = 3, nolabels = T)) +
  geom_sf(data = segmentation_smoothed %>% st_filter(sampling_square_temp) %>% st_transform(4326), color = "black", fill = "#1E88E5", alpha = 0.5) +
  theme_bw() -> f2

f1 / f2
```

\newpage
## Analysis

The three performance metrics are the following ones :
1. Symmetric mean absolute percentage error (SMAPE) of polygon density
 - The value is positive when the segmentation have a higher density than the ecoforestry polygon and negative when the segmentation have a lower density
 - In the example above, the SMAPE of the density is 0.17 as the ecoforestry have a density of 7.25 polygon/km² (1465 polygons) and the segmentation a density of 8.59 polygon/km² (1650 polygons)

2. Overlapping kernel density of the polygon stand area (left figure)
 - This performance metric represent how the distribution of the segmentation polygon stand area suit well the distribution of the ecoforestry polygon area

3. Overlapping kernel density of the polygon stand shape index (right figure)
 - The same way as the last performance metric, this performance metric represent how the distribution of the segmentation polygon stand shape index suit well the distribution of the ecoforestry polygon shape index
 - The shape index was computed with smoothed polygon to avoid overestimating the perimeter of polygons

The assumption behind this analysis is that the automated segmented polygon stand should have the same distribution of polygon area and shape than the reference ecoforestery polygon stand and therefore be operationally plausible forest stand.

```{r}
#| label = "show segmentation 2",
#| echo = FALSE
dens_polygon_forest <- nrow(polygon_forest_temp)/as.numeric(sum(st_area(polygon_forest_temp))/1000000)
dens_segmentation <- nrow(segmentation)/as.numeric(sum(st_area(segmentation))/1000000)
smape_dens <- smape(dens_polygon_forest, dens_segmentation)

ggplot() +
  geom_histogram(aes(x = segmentation_area_raster/10000,
                     y = after_stat(density)),
                 bins = 100,
                 fill = "#D81B60",
                 alpha = 0.3) +
  geom_density(aes(x = segmentation_area_raster/10000),
               color = "#D81B60",
               linewidth = 1) +
  
  geom_histogram(aes(x = polygon_forest_area_raster/10000,
                     y = after_stat(density)),
                 bins = 100,
                 fill = "#0B4D00",
                 alpha = 0.3) +
  geom_density(aes(x = polygon_forest_area_raster/10000),
               color = "#0B4D00",
               linewidth = 1) +
  labs(x = "Polygon area (ha)",
       y = "Density") +
  theme_bw() -> p_area_raster

ggplot() +
  geom_histogram(aes(x = segmentation_shape_index_raster,
                     y = after_stat(density)),
                 bins = 100,
                 fill = "#D81B60",
                 alpha = 0.3) +
  geom_density(aes(x = segmentation_shape_index_raster),
               color = "#D81B60",
               linewidth = 1) +
  
  geom_histogram(aes(x = segmentation_smoothed_shape_index_raster,
                     y = after_stat(density)),
                 bins = 100,
                 fill = "#1E88E5",
                 alpha = 0.3) +
  geom_density(aes(x = segmentation_smoothed_shape_index_raster),
               color = "#1E88E5",
               linewidth = 1) +
  
  geom_histogram(aes(x = polygon_forest_shape_index_raster,
                     y = after_stat(density)),
                 bins = 100,
                 fill = "#0B4D00",
                 alpha = 0.3) +
  geom_density(aes(x = polygon_forest_shape_index_raster),
               color = "#0B4D00",
               linewidth = 1) +
  labs(x = "Shape index",
       y = "Density") +
  theme_bw() -> p_shape_raster

p_area_raster / p_shape_raster
```

```{r}
#| label = "results",
#| eval = FALSE,
#| echo = FALSE,
#| include = FALSE
forest <- c("RMF", "OVF")
thresh <- seq(10, 100, 10)
spec <- seq(0.1, 1, 0.1)
spat <- seq(0.1, 1, 0.1)
models <- list(model_A = c("z_p95", "z_above2", "z_cv"),
               model_B = c("z_p95", "z_above2", "z_skew"),
               model_D = c("z_p95", "z_above2", "sagawi"),
               model_F = c("z_p95", "z_above2", "B6"),
               model_G = c("B6", "B8", "B11"))

models_df <- data.frame(model_name = names(models),
                        model_var  = sapply(models, paste, collapse = ", "),
                        stringsAsFactors = FALSE)

params <- merge(expand.grid(forest = forest,
                            thresh = thresh,
                            spec = spec,
                            spat = spat,
                            stringsAsFactors = FALSE),
                models_df)

# Remove done files
params %>%
  mutate(done = file.exists(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/", forest, "/", model_name, "/tables/results_thresh", thresh, "_spec", spec*100, "_spat", spat*100, ".rds"))) %>%
  filter(!done) -> params

# p <- 1

params %>%
  filter(forest == "OVF",
         model_name == "model_A",
         thresh == 50,
         spec == 0.5,
         spat == 0.5) -> param

batch_size <- ncores()  # petit !

indices <- split(seq_len(nrow(params)),
                 ceiling(seq_along(seq_len(nrow(params))) / batch_size))

for(batch in indices){

  plan(callr, workers = round(lasR::ncores()*0.25))

  future.apply::future_lapply(batch, function(p){

    tryCatch({

      library(tidyverse)
      library(sf)
      library(terra)
      library(overlapping)
      library(patchwork)
      library(rmapshaper)
      library(magrittr)

      param = params[p,]

      cat("Forest :", param$forest, "/ Model :", param$model_name, "/ Threshold :", param$thresh, "/ Spec :", param$spec, "/ Spat :", param$spat, "\n")

      if(!dir.exists(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/", param$forest))){
        dir.create(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/", param$forest))
      }

      if(!dir.exists(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/", param$forest, "/", param$model_name))){
        dir.create(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/", param$forest, "/", param$model_name))
      }

      if(!dir.exists(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/", param$forest, "/", param$model_nam, "/tables"))){
        dir.create(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/", param$forest, "/", param$model_name, "/tables"))
      }

      if(!dir.exists(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/", param$forest, "/", param$model_name, "/figures"))){
        dir.create(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/", param$forest, "/", param$model_name, "/figures"))
      }

      if(!dir.exists(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/", param$forest, "/", param$model_name, "/maps"))){
        dir.create(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/", param$forest, "/", param$model_name, "/maps"))
      }

      output_name <- paste0("_thresh", param$thresh,
                            "_spec", param$spec*100,
                            "_spat", param$spat*100)

      terra::rast(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/base_data/metrics_", param$forest, ".tif")) -> metrics

      sf::st_read(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/base_data/polygon_forest_FOR_", param$forest, ".shp"), quiet = T) %>%
        dplyr::select() %>%
        sf::st_cast("MULTIPOLYGON") %>%
        sf::st_cast("POLYGON") %>%
        mutate(area = as.numeric(sf::st_area(.)),
               perimeter = as.numeric(sf::st_perimeter(.)),
               perimeter_area_ratio = perimeter/area,
               shape_index = perimeter / sqrt(pi * area)) %>%
        filter_if(is.numeric, is.finite) -> polygon_forest

      sf::st_read(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/segmentation/", param$forest, "/", param$model_name, "/segmentation", output_name, ".shp"), quiet = T) %>%
        sf::st_cast("MULTIPOLYGON") %>%
        sf::st_cast("POLYGON") %>%
        mutate(area = as.numeric(st_area(.)),
               perimeter = as.numeric(st_perimeter(.)),
               perimeter_area_ratio = perimeter/area,
               shape_index = perimeter / sqrt(pi * area)) %>%
        filter_if(is.numeric, is.finite) -> segmentation

      sf::st_read(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/segmentation/", param$forest, "/", param$model_name, "/segmentation", output_name, "_smoothed.shp"), quiet = T) %>%
        sf::st_make_valid() %>%
        sf::st_cast("MULTIPOLYGON") %>%
        sf::st_cast("POLYGON") %>%
        mutate(area = as.numeric(st_area(.)),
               perimeter = as.numeric(st_perimeter(.)),
               perimeter_area_ratio = perimeter/area,
               shape_index = perimeter / sqrt(pi * area)) %>%
        filter_if(is.numeric, is.finite) -> segmentation_smoothed

      polygon_forest %>%
        dplyr::select(area) %>%
        terra::rasterize(metrics, field = "area") %>%
        terra::values(na.rm = TRUE) -> polygon_forest_area_raster

      polygon_forest_area_raster[is.finite(polygon_forest_area_raster)] -> polygon_forest_area_raster

      segmentation %>%
        dplyr::select(area) %>%
        terra::rasterize(metrics, field = "area") %>%
        terra::values(na.rm = TRUE) -> segmentation_area_raster

      segmentation_area_raster[is.finite(segmentation_area_raster)] -> segmentation_area_raster

      overlap_area_raster_OV <- overlapping::overlap(list(segmentation_area_raster, polygon_forest_area_raster))$OV

      # ggplot() +
      #   geom_density(data = segmentation, aes(x = area/10000), color = "#D81B60") +
      #   geom_density(data = polygon_forest, aes(x = area/10000), color = "#0B4D00") +
      #   theme_bw() -> p_area_poly

      ggplot() +
        geom_histogram(data = segmentation %>% dplyr::filter(is.finite(area)),
                       aes(x = area/10000,
                           y = after_stat(density)),
                       bins = 100,
                       fill = "#D81B60",
                       alpha = 0.3) +
        geom_density(data = segmentation %>% dplyr::filter(is.finite(area)),
                     aes(x = area/10000),
                     color = "#D81B60",
                     linewidth = 1) +

        geom_histogram(data = polygon_forest %>% dplyr::filter(is.finite(area)),
                       aes(x = area/10000,
                           y = after_stat(density)),
                       bins = 100,
                       fill = "#0B4D00",
                       alpha = 0.3) +
        geom_density(data = polygon_forest %>% dplyr::filter(is.finite(area)),
                     aes(x = area/10000),
                     color = "#0B4D00",
                     linewidth = 1) +
        theme_bw() -> p_area_poly

      # ggplot() +
      #   geom_density(data = segmentation_area_raster, aes(x = area/10000), color = "#D81B60") +
      #   geom_density(data = polygon_forest_area_raster, aes(x = area/10000), color = "#0B4D00") +
      #   theme_bw() -> p_area_raster

      ggplot() +
        geom_histogram(aes(x = segmentation_area_raster/10000,
                           y = after_stat(density)),
                       bins = 100,
                       fill = "#D81B60",
                       alpha = 0.3) +
        geom_density(aes(x = segmentation_area_raster/10000),
                     color = "#D81B60",
                     linewidth = 1) +

        geom_histogram(aes(x = polygon_forest_area_raster/10000,
                           y = after_stat(density)),
                       bins = 100,
                       fill = "#0B4D00",
                       alpha = 0.3) +
        geom_density(aes(x = polygon_forest_area_raster/10000),
                     color = "#0B4D00",
                     linewidth = 1) +
        theme_bw() -> p_area_raster

      p_area_poly / p_area_raster -> p_area

      ggsave(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/", param$forest, "/", param$model_name, "/figures/area", output_name, ".jpg"),
             plot = p_area,
             width = 15,
             height = 10,
             dpi = 150)

      graphics.off()
      rm(polygon_forest_area_raster,
         segmentation_area_raster,
         p_area_poly,
         p_area_raster,
         p_area)
      gc()

      polygon_forest %>%
        dplyr::select(shape_index) %>%
        terra::rasterize(metrics, field = "shape_index") %>%
        terra::values(na.rm = TRUE) -> polygon_forest_shape_index_raster

      polygon_forest_shape_index_raster[is.finite(polygon_forest_shape_index_raster)] -> polygon_forest_shape_index_raster

      segmentation %>%
        dplyr::select(shape_index) %>%
        terra::rasterize(metrics, field = "shape_index") %>%
        terra::values(na.rm = TRUE) -> segmentation_shape_index_raster

      segmentation_shape_index_raster[is.finite(segmentation_shape_index_raster)] -> segmentation_shape_index_raster

      segmentation_smoothed %>%
        dplyr::select(shape_index) %>%
        terra::rasterize(metrics, field = "shape_index") %>%
        terra::values(na.rm = TRUE) -> segmentation_smoothed_shape_index_raster

      segmentation_smoothed_shape_index_raster[is.finite(segmentation_smoothed_shape_index_raster)] -> segmentation_smoothed_shape_index_raster

      overlap_shape_raster_OV <- overlapping::overlap(list(segmentation_shape_index_raster, polygon_forest_shape_index_raster))$OV

      overlap_shape_smoothed_raster_OV <- overlapping::overlap(list(segmentation_smoothed_shape_index_raster, polygon_forest_shape_index_raster))$OV

      # ggplot() +
      #   geom_density(data = segmentation, aes(x = shape_index), color = "#D81B60") +
      #   geom_density(data = segmentation_smoothed, aes(x = shape_index), color = "#1E88E5") +
      #   geom_density(data = polygon_forest, aes(x = shape_index), color = "#0B4D00") +
      #   theme_void() -> p_shape_poly

      ggplot() +
        geom_histogram(data = segmentation %>% dplyr::filter(is.finite(shape_index)),
                       aes(x = shape_index,
                           y = after_stat(density)),
                       bins = 100,
                       fill = "#D81B60",
                       alpha = 0.3) +
        geom_density(data = segmentation %>% dplyr::filter(is.finite(shape_index)),
                     aes(x = shape_index),
                     color = "#D81B60",
                     linewidth = 1) +

        geom_histogram(data = segmentation_smoothed %>% dplyr::filter(is.finite(shape_index)),
                       aes(x = shape_index,
                           y = after_stat(density)),
                       bins = 100,
                       fill = "#1E88E5",
                       alpha = 0.3) +
        geom_density(data = segmentation_smoothed %>% dplyr::filter(is.finite(shape_index)),
                     aes(x = shape_index),
                     color = "#1E88E5",
                     linewidth = 1) +

        geom_histogram(data = polygon_forest %>% dplyr::filter(is.finite(shape_index)),
                       aes(x = shape_index,
                           y = after_stat(density)),
                       bins = 100,
                       fill = "#0B4D00",
                       alpha = 0.3) +
        geom_density(data = polygon_forest %>% dplyr::filter(is.finite(shape_index)),
                     aes(x = shape_index),
                     color = "#0B4D00",
                     linewidth = 1) +
        theme_bw() -> p_shape_poly

      # ggplot() +
      #   geom_density(data = segmentation_shape_index_raster, aes(x = shape_index), color = "#D81B60") +
      #   geom_density(data = segmentation_smoothed_shape_index_raster, aes(x = shape_index), color = "#1E88E5") +
      #   geom_density(data = polygon_forest_shape_index_raster, aes(x = shape_index), color = "#0B4D00") +
      #   theme_void() -> p_shape_raster

      ggplot() +
        geom_histogram(aes(x = segmentation_shape_index_raster,
                           y = after_stat(density)),
                       bins = 100,
                       fill = "#D81B60",
                       alpha = 0.3) +
        geom_density(aes(x = segmentation_shape_index_raster),
                     color = "#D81B60",
                     linewidth = 1) +

        geom_histogram(aes(x = segmentation_smoothed_shape_index_raster,
                           y = after_stat(density)),
                       bins = 100,
                       fill = "#1E88E5",
                       alpha = 0.3) +
        geom_density(aes(x = segmentation_smoothed_shape_index_raster),
                     color = "#1E88E5",
                     linewidth = 1) +

        geom_histogram(aes(x = polygon_forest_shape_index_raster,
                           y = after_stat(density)),
                       bins = 100,
                       fill = "#0B4D00",
                       alpha = 0.3) +
        geom_density(aes(x = polygon_forest_shape_index_raster),
                     color = "#0B4D00",
                     linewidth = 1) +
        theme_bw() -> p_shape_raster

      p_shape_poly / p_shape_raster -> p_shape

      ggsave(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/", param$forest, "/", param$model_name, "/figures/shape", output_name, ".jpg"),
             plot = p_shape,
             width = 15,
             height = 10,
             dpi = 150)

      graphics.off()
      rm(polygon_forest_shape_index_raster,
         segmentation_shape_index_raster,
         segmentation_smoothed_shape_index_raster,
         p_shape_poly,
         p_shape_raster,
         p_shape)
      gc()

      polygon_forest %>%
        slice(1000) %>%
        sf::st_centroid() %>%
        sf::st_buffer(1000) -> clip_area

      ggplot() +
        # base_map(clip_area %>% st_transform(4326) %>% st_bbox, basemap = "google-satellite", increase_zoom = 3, nolabels = T) +
        geom_sf(data = polygon_forest %>% st_filter(clip_area) %>% rmapshaper::ms_clip(clip_area) %>% sf::st_transform(4326), color = "#D81B60", fill = NA, lwd = 1) +
        geom_sf(data = segmentation_smoothed %>% st_filter(clip_area) %>% rmapshaper::ms_clip(clip_area) %>% sf::st_transform(4326), color = "#1E88E5", fill = NA, lwd = 1) +
        theme_bw() -> p_map

      ggsave(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/", param$forest, "/", param$model_name, "/maps/map", output_name, ".jpg"),
             plot = p_map,
             width = 10,
             height = 10,
             dpi = 300)

      graphics.off()
      rm(p_map,
         clip_area)
      gc()

      tibble(forest = param$forest,
             model_name = param$model_name,
             thresh = param$thresh,
             spec = param$spec,
             spat = param$spat,

             overlap_area = overlapping::overlap(list(segmentation$area, polygon_forest$area))$OV,
             overlap_area_raster = overlap_area_raster_OV,

             overlap_shape = overlapping::overlap(list(segmentation$shape_index, polygon_forest$shape_index))$OV,
             overlap_shape_raster = overlap_shape_raster_OV,

             overlap_shape_smoothed = overlapping::overlap(list(segmentation_smoothed$shape_index, polygon_forest$shape_index))$OV,
             overlap_shape_smoothed_raster = overlap_shape_smoothed_raster_OV,

             mean_area_forest = mean(polygon_forest$area),
             mean_area_segmentation = mean(segmentation$area),
             mean_area_smape = smape(mean_area_forest, mean_area_segmentation),

             sd_area_forest = sd(polygon_forest$area),
             sd_area_segmentation = sd(segmentation$area),
             sd_area_smape = smape(sd_area_forest, sd_area_segmentation),

             n_forest = nrow(polygon_forest),
             n_segmentation = nrow(segmentation),
             n_smape = smape(n_forest, n_segmentation),

             dens_forest = n_forest/(as.numeric(sum(st_area(polygon_forest)))/1000000),
             dens_segmentation = n_segmentation/(as.numeric(sum(st_area(segmentation)))/1000000),
             dens_smape = smape(dens_forest, dens_segmentation)) %>%

        bind_cols(quantile(polygon_forest$area,
                           probs = seq(0, 1, 0.1)) %>%
                    as.list() %>%
                    tibble::as_tibble_row(.name_repair = "minimal") %>%
                    rename_with(~ paste0("quantile_forest_", seq(0, 100, 10))),
                  quantile(segmentation$area,
                           probs = seq(0, 1, 0.1)) %>%
                    as.list() %>%
                    tibble::as_tibble_row(.name_repair = "minimal") %>%
                    rename_with(~ paste0("quantile_segmentation_", seq(0, 100, 10)))) -> results

      results %>%
        saveRDS(paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/", param$forest, "/", param$model_name, "/tables/results", output_name, ".rds"))

      terra::tmpFiles(remove = TRUE)
      rm(metrics,
         polygon_forest,
         segmentation,
         chunk_size,
         segmentation_chunks,
         segmentation_smoothed,
         overlap_area_raster_OV,
         overlap_shape_raster_OV,
         overlap_shape_smoothed_raster_OV)
      gc()

      return(NULL)

    }, error = function(e) {

      cat(c("Error at index ", p, " : ", conditionMessage(e), "----------"),
          file = paste0("D:/00_Ontario_eFRI/segmentation/grm_parametrization/log_error/log_error_", param$forest, "_", param$model_name, "_thresh", param$thresh, "_spec", param$spec*100, "_spat", param$spat*100, ".txt"),
          append = TRUE,
          sep = "\n")

      return(NULL)
    })

  })

}

list.files("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results", full.names = TRUE) %>%
  map(list.files, full.names = TRUE) %>%
  unlist() %>%
  map(list.files, full.names = TRUE, pattern = "tables") %>%
  unlist() %>%
  list.files(full.names = TRUE, pattern = "\\.rds$") %T>%
  {files_names <<- .} %>%
  map_dfr(readRDS) -> results

# Forgot to save the model name in th first round, but its now corrected at line 721, so I need to add it manually for the first results
files_names %>%
  str_extract("model_[A-Z]") -> model_names

results %>%
  mutate(model_name = model_names) %>%
  select(forest, model_name, everything()) -> results

results %>%
  saveRDS("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/results.rds")
```

```{r}
#| label = "figures and tables",
#| eval = FALSE,
#| echo = FALSE,
#| include = FALSE
# Parameters
dens_threshold <- 0.1
weight_area <- 1/2
weight_shape <- 1/2

# Read data
readRDS("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/results.rds") %>%
  mutate(area_score = overlap_area_raster,
         shape_score = overlap_shape_smoothed_raster) %>%
  mutate(final_score = (weight_area*(1-area_score)) + (weight_shape*(1-shape_score))) %>%
  mutate_at(c("thresh", "spec", "spat"), as.numeric) -> results

# Compute for a specific forest and model and produce figures
selected_forest <- "RMF"
selected_forest <- "OVF"
selected_model <- "model_B"

results %>%
  filter(forest == selected_forest) %>%
  filter(model_name == selected_model) %>%
  filter(dens_smape >= -dens_threshold & dens_smape <= dens_threshold) %>%
  mutate(parameters = paste0("thresh", thresh, "_spec", spec*100, "_spat", spat*100)) -> results_temp

#plot_ly(data = results_temp %>% filter(spec >= 50),
plot_ly(data = results_temp,
        x = ~area_score, y = ~shape_score, z = ~dens_smape,
        color = ~final_score,
        colors = c("#117733", "#882235"),
        #size = ~thresh,
        #size = ~spat,
        size = ~spec,
        sizes = c(1, 1000),
        type = "scatter3d",
        mode = "markers",
        text = ~parameters,
        hovertemplate = paste("Overlap Area : %{x}<br>",
                              "Overlap shape : %{y}<br>",
                              "Symmetric Mean Absolute Percentage Error (SMAPE) : %{z}<br>",
                              "Parameters : %{text}<extra></extra>"))

plot_ly(data = results_temp,
        x = ~thresh, y = ~spec, z = ~spat,
        color = ~ dens_smape,
        colors = c("#332288", "#117733", "#882235"),
        size = ~ dens_smape,
        sizes = c(1, 1000),
        type = "scatter3d",
        mode = "markers",
        hovertemplate = paste("Threshold : %{x}<br>",
                              "Spec : %{y}<br>",
                              "Spat : %{z}<br>",
                              "Symmetric Mean Asbolute Error (SMAPE) : %{marker.color}<extra></extra>"))

plot_ly(data = results_temp,
        x = ~thresh, y = ~spec, z = ~spat,
        color = ~area_score,
        colors = c("#882235", "#117733"),
        size = ~area_score,
        sizes = c(1, 1000),
        type = "scatter3d",
        mode = "markers",
        hovertemplate = paste("Threshold : %{x}<br>",
                              "Spec : %{y}<br>",
                              "Spat : %{z}<br>",
                              "Overlap Area : %{marker.color}<extra></extra>"))

plot_ly(data = results_temp,
        x = ~thresh, y = ~spec, z = ~spat,
        color = ~shape_score,
        colors = c("#882235", "#117733"),
        size = ~shape_score,
        sizes = c(1, 1000),
        type = "scatter3d",
        mode = "markers",
        hovertemplate = paste("Threshold : %{x}<br>",
                              "Spec : %{y}<br>",
                              "Spat : %{z}<br>",
                              "Overlap Shape : %{marker.color}<extra></extra>"))

plot_ly(data = results_temp,
        x = ~thresh, y = ~spec, z = ~spat,
        color = ~final_score,
        colors = c("#117733", "#882235"),
        size = ~final_score,
        sizes = c(1000, 1),
        type = "scatter3d",
        mode = "markers",
        hovertemplate = paste("Threshold : %{x}<br>",
                              "Spec : %{y}<br>",
                              "Spat : %{z}<br>",
                              "Final score : %{marker.color}<extra></extra>"))

results_temp %>%
  arrange(final_score) %>%
  slice(1) %>%
  dplyr::select(forest, model_name, thresh, spec, spat, overlap_area_raster, overlap_shape_smoothed_raster, n_forest, n_segmentation, dens_smape, final_score) -> best_parameters_temp

# Compute for each forest and model
results %>%
  group_by(forest, model_name) %>%
  reframe() -> forest_model_combinations

map_dfr(seq_len(nrow(forest_model_combinations)),
        function(x){

          results %>%
            filter(forest == forest_model_combinations$forest[x]) %>%
            filter(model_name == forest_model_combinations$model_name[x]) %>%
            filter(dens_smape >= -dens_threshold & dens_smape <= dens_threshold) %>%
            # Median values
            # slice_min(abs(spec - median(spec))) %>%
            # slice_min(abs(spat - median(spat))) %>%
            # slice_min(abs(thresh - median(thresh))) %>%
            # Min final score
            arrange(final_score) %>%
            slice(1) %>%
            # Final selection
            dplyr::select(forest, model_name, thresh, spec, spat, overlap_area_raster, overlap_shape_smoothed_raster, n_forest, n_segmentation, dens_smape, final_score)

        }) %>%
  mutate(type = "by_forest_model") -> best_parameters_by_forest_model

# Compute for each forest
best_parameters_by_forest <- results %>%
  group_by(forest, thresh, spec, spat) %>%
  reframe(area_score = mean(area_score),
          shape_score = mean(shape_score),
          dens_smape = mean(dens_smape)) %>%
  mutate(final_score = (weight_area*(1-area_score)) + (weight_shape*(1-shape_score))) %>%
  filter(dens_smape >= -dens_threshold & dens_smape <= dens_threshold) %>%
  group_by(forest) %>%
  # # Median values
  # slice_min(abs(spec - median(spec))) %>%
  # slice_min(abs(spat - median(spat))) %>%
  # slice_min(abs(thresh - median(thresh)))
  # Min final score
  arrange(final_score) %>%
  slice(1)

map_dfr(unique(results$forest),
        function(f){

          best_parameters_by_forest %>%
            filter(forest == f) -> best_parameters_by_forest_temp

          results %>%
            filter(forest == f) %>%
            filter(thresh == best_parameters_by_forest_temp$thresh &
                     spec == best_parameters_by_forest_temp$spec &
                     spat == best_parameters_by_forest_temp$spat) %>%
            dplyr::select(forest, model_name, thresh, spec, spat, overlap_area_raster, overlap_shape_smoothed_raster, n_forest, n_segmentation, dens_smape, final_score)

        }) %>%
  mutate(type = "by_forest") -> best_parameters_by_forest

# Compute general results
best_parameters <- results %>%
  group_by(thresh, spec, spat) %>%
  reframe(area_score = mean(area_score),
          shape_score = mean(shape_score),
          dens_smape = mean(dens_smape)) %>%
  mutate(final_score = (weight_area*(1-area_score)) + (weight_shape*(1-shape_score))) %>%
  filter(dens_smape >= -dens_threshold & dens_smape <= dens_threshold) %>%
  # # Median values
  # slice_min(abs(spec - median(spec))) %>%
  # slice_min(abs(spat - median(spat))) %>%
  # slice_min(abs(thresh - median(thresh)))
  # # Min final score
  arrange(final_score) %>%
  slice(1)

results %>%
  filter(thresh == best_parameters$thresh &
           spec == best_parameters$spec &
           spat == best_parameters$spat) %>%
  dplyr::select(forest, model_name, thresh, spec, spat, overlap_area_raster, overlap_shape_smoothed_raster, n_forest, n_segmentation, dens_smape, final_score) %>%
  mutate(type = "all") -> best_parameters

# Merge all results
bind_rows(best_parameters_by_forest_model,
          best_parameters_by_forest,
          best_parameters) -> best_parameters_all

best_parameters_all %>%
  saveRDS("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/best_parameters_all.rds")
```

\newpage
## Results

The results presented here are, as mentionned before, in the Ottawa Valley Forest for the model_A (z_p95 + z_above2 + z_cv). All the follow process can be made for every forest and every combinaison of metrics. The threshold of the SMAPE is 0.1 which represent more of less 10 % of polygon density variation. All sets of parameters that gave SMAPE lower than -0.1 and higher than 0.1 have been removed. 

```{r}
#| label = "show figures and tables 1",
#| echo = FALSE
# Parameters
dens_threshold <- 0.1
weight_area <- 1/2
weight_shape <- 1/2

# Read data
readRDS("D:/00_Ontario_eFRI/segmentation/grm_parametrization/results/results.rds") %>%
  rename(dens_smape = dens_sspe) %>% 
  mutate(area_score = overlap_area_raster,
         shape_score = overlap_shape_smoothed_raster) %>%
  mutate(final_score = (weight_area*(1-area_score)) + (weight_shape*(1-shape_score))) %>%
  mutate_at(c("thresh", "spec", "spat"), as.numeric) -> results

# Compute for a specific forest and model and produce figures
selected_forest <- "OVF"
selected_model <- "model_A"

results %>%
  filter(forest == selected_forest) %>%
  filter(model_name == selected_model) %>%
  filter(dens_smape >= -dens_threshold & dens_smape <= dens_threshold) %>%
  mutate(parameters = paste0("thresh", thresh, "_spec", spec*100, "_spat", spat*100)) -> results_temp

# Color scaling
pal <- colorRampPalette(c("#332288", "#117733", "#882235"))
col_scaled <- pal(100)[cut(results_temp$dens_smape, breaks = 100)]

# Size scaling
cex_scaled <- scales::rescale(results_temp$dens_smape, to = c(0.5, 2))

# Plot
s3d <- scatterplot3d(
  x = results_temp$thresh,
  y = results_temp$spec,
  z = results_temp$spat,
  color = col_scaled,
  pch = 16,
  cex.symbols = cex_scaled,
  xlab = "Threshold",
  ylab = "Spec",
  zlab = "Spat"
)
```

\newpage
After this step, the optimal parameters was choosen as the one that maximized both the overlap of the areas and the shape index.

```{r}
#| label = "show figures and tables 2",
#| echo = FALSE
# Color scaling
pal <- colorRampPalette(c("#882235", "#117733"))
col_scaled <- pal(100)[cut(results_temp$area_score, breaks = 100)]

# Size scaling
cex_scaled <- scales::rescale(results_temp$area_score, to = c(0.5, 2))

# Plot
s3d_area <- scatterplot3d(
  x = results_temp$thresh,
  y = results_temp$spec,
  z = results_temp$spat,
  color = col_scaled,
  pch = 16,
  cex.symbols = cex_scaled,
  xlab = "Threshold",
  ylab = "Spec",
  zlab = "Spat"
)

# Color scaling
pal <- colorRampPalette(c("#882235", "#117733"))
col_scaled <- pal(100)[cut(results_temp$shape_score, breaks = 100)]

# Size scaling
cex_scaled <- scales::rescale(results_temp$shape_score, to = c(0.5, 2))

# Plot
s3d_shape <- scatterplot3d(
  x = results_temp$thresh,
  y = results_temp$spec,
  z = results_temp$spat,
  color = col_scaled,
  pch = 16,
  cex.symbols = cex_scaled,
  xlab = "Threshold",
  ylab = "Spec",
  zlab = "Spat"
)
```

\newpage
The final score was computed as follow :
- final_score = (0.5 x (1-overlap_area)) + (0.5 x (1-overlap_shape_index))
- the best parameters set was choosen as the lowest final score among all combinaison that had a SMAPE of polygon density of more of less 0.1.
```{r}
#| label = "show figures and tables 3",
#| echo = FALSE
results_temp %>%
  arrange(final_score) %>%
  slice(1) %>%
  dplyr::select(forest, model_name, thresh, spec, spat, overlap_area_raster, overlap_shape_smoothed_raster, n_forest, n_segmentation, dens_smape, final_score) -> best_parameters_temp

local({

# Color scaling
pal <- colorRampPalette(c("#117733", "#882235"))
col_scaled <- pal(100)[cut(results_temp$final_score, breaks = 100)]

# Size scaling
cex_scaled <- scales::rescale(results_temp$final_score, to = c(2, 0.5))

# Plot
s3d <- scatterplot3d(
  x = results_temp$thresh,
  y = results_temp$spec,
  z = results_temp$spat,
  color = col_scaled,
  pch = 16,
  cex.symbols = cex_scaled,
  xlab = "Threshold",
  ylab = "Spec",
  zlab = "Spat"
)

i <- which.min(results_temp$final_score)

coords <- s3d$xyz.convert(
  results_temp$thresh[i],
  results_temp$spec[i],
  results_temp$spat[i]
)

points(coords$x, coords$y, col = "red", pch = 19, cex = 2)

})


```

\newpage
## References

